# Methods

## Participants

We used data from an observational study of brain infection conducted at the Hospital for Tropical Diseases (HTD), Ho Chi Minh City, Vietnam, a large centre providing secondary and tertiary treatment for a wide range of tropical infections[@thwaites2002]. This study received ethical approvals from HTD and the Oxford Tropical Research Ethics Committee[@donovan2020]. Participants were enrolled between 29th August 2017 and 22nd January 2021. Inclusion criteria included a minimum age of 16 years, having suspected brain infection, admitted to a neuro-infection ward, and undergoing lumbar puncture at baseline as a routine diagnostic procedure. Patients were ineligible for enrolment if performing a lumbar puncture was contraindicated, or if informed consent to join the study was not given (by the patient, or by a relative if the patient lacked capacity to consent). Specifically for this analysis, we also excluded patients with contaminated mycobacterial culturing results and with no subsequent culture in the first week since admission, concepts of which were later discussed in \@ref(data-collection-and-testing procedures).

## Data collection and testing procedures

For all participants, demographic data and medical history were recorded. HIV testing was performed on a case-by-case basis by the treating clinician. Patients admitted with a condition suggestive of immune deficiency or had been involved in high risk activities were tested for HIV. Patients with either a known HIV positive status, or a positive HIV test result, were considered HIV positive. Patients underwent clinical examination and laboratorial investigations according to main study protocol, including: blood test, sputum, and lumbar puncture, unless contra-indicated. Cerebro-spinal fluid (CSF) was obtained by lumbar puncture, with CSF white blood cells and cellular differential, CSF protein, CSF glucose (with paired blood glucose), and CSF lactate routinely measured. Optimally 6mls CSF was used for mycobacterial testing unless less CSF was available (in which case mycobacterial testing was still performed). Confirmatory tests for TBM performed on CSF and used for this analysis were Ziehl--Neelsen staining and smear (ZN Smear), mycobacterial culture using Mycobacteria Growth Indicator Tube (MGIT), and Xpert or Xpert MTB/RIF Ultra (XpertUltra). Since Xpert and XpertUltra were found diagnostically comparable in a subset of these data[@donovan2020], Xpert and XpertUltra were considered the same in this analysis and, are both denoted as Xpert. Confirmatory mycobacterial testing was performed if TBM was suspected by the treating clinician.

Methods of CSF processing have been described elsewhere[@nhu2013, @donovan2020]. Briefly: CSF samples were centrifuged at 3000g for 15 minutes, and most of the CSF supernatant was removed. The CSF deposit was resuspended in 500µL of remaining supernatant, with this resuspended pellet then used for ZN smear (100µL), MGIT (200µL), and either Xpert or XpertUltra (200 µL).

All patients received appropriate anti-TB chemotherapy regimens according to national and local treatment guidelines, depending on the diagnosis, without any interference from the study. At the time of discharge or death, all patients received a final diagnosis. If at least one of ZN Smear, Xpert, Xpert Ultra, or MGIT was positive at any time during the follow-up, the patient was considered confirmed TBM; otherwise, if TBM was clinically suspected and treated, with confirmatory microbiological tests negative, the patient would be considered suspected TBM. If the patient recovered without anti-TB chemotherapy they was reassigned to another diagnosis (i.e. not TBM).

In case of repeated CSF sampling (performed based on clinical need), only the first sample with at least 3mls of CSF collected not later than the first week since admission were used. Results of haematological and biochemical parameters, and of ZN Smear, Xpert, Xpert Ultra, and MGIT, were taken from the same CSF sample (in addition to the paired blood glucose).

## Statistical analysis

### Latent class regression model

We conducted a Bayesian Latent Class Model in which the three aforementioned TBM confirmatory tests (ZN Smear, MGIT, and Xpert) were used as manifest variables and the target latent class is the actual disease status (TBM or not TBM). All three tests are detection tool via direct visualisation under microscope (ZN Smear), via their DNA (Xpert), or via grown culture (MGIT). Predictors for the latent class were chosen according to background knowledge of potential risk factors [@marais2010] and are summarised in table \@ref(tab:predictor-tab). We added three parameters: (1) CSF oeosinophil count - a strong bio-marker for oeosinophilic meningitis, a condition usually caused by parasitic helminths; (2) Cryptococcus Antigen/Indian Ink - a confirmatory test for parasitic infection; and (3) CSF erythrocyte count - an independent marker for traumatic lumbar puncture, in which case we also made a correction to white cell counts and biochemical features[@greenberg2008; @nigrovic2011; @mehl1986]. Details of the model design are summarised in the supplementary document.

```{r predictor-tab, out.width="100%",  tab.cap="Contribution of different features to TBM prevalence and patients' mycobacillary burden for each model mentioned in table \\@ref(tab:model-archs), based on background knowledge. Cell values are the expected direction of association and level of confidence; empty cells mean no association assumed", tab.id = "predictor-tab", label='tab0'}
tibble::tribble(
  ~ 'Predictor'                    , ~ 'TBM prevalence', ~ 'Bacillary Burden',
  'Age'                            , '+, weak'  , ''                  ,
  'HIV infection'                  , '+, strong', '+, weak'    ,
  'Past TB contact'                , '+, weak'  , ''                  ,
  'TB-suggested symptoms'          , '+, weak'  , ''                  ,
  'Local motor deficit'            , '+, weak'  , ''                  ,         
  'Cranial nerve palsy'            , '+, weak'  , ''                  ,
  'Days from onset'                , '+, weak'  , ''                  ,
  'PTB/X-Ray'                      , '+, weak'  , ''                  ,
  'MTB/X-Ray'                      , '+, strong', ''                  ,
  'Glasgow Coma Score'             , '-, weak'  , ''                  ,
  'Cryptococcus Antigen/Indian Ink', '-, weak'  , ''                  ,
  'Blood Glucose'                  , '-, weak'  , ''                  ,
  'CSF Glucose'                    , '-, weak'  , '?, weak'     ,
  'CSF Lymphocyte Count'           , '+, weak'  , '-, weak'    ,
  'CSF Total While cell Count'     , '+-^[Risk of TBM peaks with intermediate CSF white cell count], weak' , '+, weak'    ,
  'CSF Protein'                    , '+, weak'  , '+, weak'    ,
  'CSF Lactate'                    , '+, weak'  , '+, weak'    ,
  'CSF Oeosinophil Count'          , '-, strong', ''                  ,
  'CSF RBC Count'                  , '?, weak'   , ''                 
) |>
  flextable::flextable() |>
  ftExtra::colformat_md(2) |>
  # flextable::footnote(i=16, j=2, 
    # value=flextable::as_paragraph('')) |>
  flextable::width(j=1, width=2) |>
  flextable::width(j=2:3, width=1.2) |>
  flextable::theme_zebra() |>
  flextable::bold(bold = FALSE, part = "footer") |>
  flextable::italic(italic = TRUE, part = "footer" )
```

```{aligned_formula, fml.lab='eq:fml-gcs', results='asis'}
\begin{aligned}
RGCS &= 15 - GCS \\
RGCS_{Eyes} &= 4 - GCS_{Eyes} \\
RGCS_{Verbal} &= 5 - GCS_{Verbal} \\
RGCS_{Motor} &= 6 - GCS_{Motor}
\end{aligned}
```

As most continuous variables were right-skewed, they were first transformed to logarithmic scale, after which they were centred and divided by corresponding standard deviations. Glasgow Coma Score (GCS) and Glasgow Coma Scale components (Voice - GCSV, Eyes - GCSE, and Muscle - GCSM) were translated to *Reversed GCS* as demonstrated in formula `r my_fn$labEq.docx('fml-gcs')`, so that a GCS of 15 is equal to a RGCS of 0 and a GCS of 3 is equal to RGCS of 12. Binary variables were dummy-coded into 0 for "Negative"/"No" and 1 for "Positive"/"Yes".

```{r gcs-tab, out.width='100%', tab.id='gcs-tab', tab.cap='Conversion table from clasic Glasgow Coma Score (GCS) to Reversed GCS (RGCS)', include=FALSE}

tibble::tribble(
  ~ Feature,                     ~ Response, ~ GCS, ~ RGCS,
  'Eye response',        'Open sponatenously', 4, 0,
  'Eye response',     'Open to voice command', 3, 1,
  'Eye response',              'Open to pain', 2, 2,
  'Eye response',               'No eye open', 1, 3,
  'Verbal response',             'Orientated', 5, 0,
  'Verbal response',               'Confused', 4, 1,
  'Verbal response',     'Inappopriate words', 3, 2,
  'Verbal response','Incomprehensible sounds', 2, 3,
  'Verbal response',     'No verbal response', 1, 4,
  'Motor response',            'Obey command', 6, 0,
  'Motor response',         'Localising pain', 5, 1,
  'Motor response',    'Withdrawal from pain', 4, 2,
  'Motor response',                 'Flexing', 3, 3,
  'Motor response',               'Extending', 2, 4,
  'Motor response',       'No motor response', 1, 5
) |>
  flextable::flextable() |>
  flextable::merge_v(j = 1) |>  
  flextable::width(j=1, width=2) |>
  flextable::width(j=2, width=2)
```

Compared to the uniform case definition, we also made a change to *Past TB contact*. We hypothesised that by interpreting the original question "contact with TB patients within the past year" into "*noticeable* contact with TB patients within the past year", we would be able to judge all "Unknown" answer as a "No". No brain imaging was taken in our data.

```{r model-archs, tab.cap="Model architectures and extensions", tab.id="model-archs"}
model_archs <-
  data.frame(
    Model=1:5,
    Def=c(
      'No bacillary burden; everyone in the same class has equal risk of tested positive',
      'Added individual bacillary burden; impacts of bacillary burden on test results are the same',
      'Impacts of bacillary burden are different for different tests',
      'Added technical fluctuation as a second random effect; fixed effects only contributes to bacillary burden',
      'Added fixed effects for technical fluctuation'
    )
  ) 

model_archs |> 
  flextable::flextable() |>
  flextable::set_header_labels(
    Model='Model', 
    Def='Base definition (Only added effects compared to lower number are mentioned)'
  ) |>
  flextable::width(j=1, width=.5)|>
  flextable::width(j=2, width=5) |>
  flextable::theme_zebra()
```

The basic design of our model is shown in figure \@ref(fig:skeleton-model). Our models consist of one logistic regression estimating the prevalence of TBM in the population (prevalence model) and linear regression estimating the latent bacilliary burden among TBM positive patients (bacillary burden model); this bacillary burden contributes to We considered a range of models increasingly flexible and less constraints[@schumacher2016] (table \@ref(tab:model-archs)). As supported by background knowledge[@marais2010], only CSF White Cell count was modeled via a quadratic effect.

```{r skeleton-model, fig.show='hold', out.width="100%", fig.align="center", fig.cap="Model basic design. Bacillary burden is only included in model 2 +. Test positive probabilities are for demonstration only and do not correspond to the real ones.", fig.id='skeleton-model'}

fig_svg <- cowplot::ggdraw() + 
  cowplot::draw_image(magick::image_read_svg("includes/classicLCA.svg", width=212*5, height=159*5))
plot(fig_svg)
```

By protocol, most patients with a very high chance of or evidently diagnosed with different diseases were not tested with the TBM confirmation assays. We assumed these patients are all negative, because these tests are known to have high specificity[@nhu2013, @heemskerk2018]. Missing predictors were handled case by case basing on their expected missing mechanisms (table \@ref(tab:missing-handling)). In case they were expected to be Missing At Random (MAR), we performed a model-based imputation.

We used a Bayesian approach for estimation. For TBM confirmatory tests sensitivity and specificity we chose Normal prior distributions, on the logit scale, with appropriate means and standard deviations, so that they covered past estimates from earlier studies involving Vietnamese cohorts[@thwaites2004; @nhu2013; @heemskerk2018] (figure \@ref(fig:mv-priors)).<!-- Weakly informative priors for True Positive Rate (TPC = Sensitivity) given the discrepancies between different research.--> We also used *Normal* distribution for the relation of covariates in both prevalance and bacilliary models, but with weakly informative configurations. Posterior distributions were obtained via Hamiltonian Markov Chain Monte Carlo, using the Stan program and its R interface in the RStan[@stan-doc] package. 

All model performances were estimated and compared by (1) expected log point-wise predictive density (elpd) [@vehtari2016], (2) calibration metrics and calibration curves[@rms; @VanCalster2019], and (3) Receiver Operating Characteristic (ROC) curves and corresponding Area Under the Curve (AUC). All metrics were estimated using a pooled dataset held out from 5 repetitions of 20-fold cross validation.<!--Pair-wise residual correlation were plotted to validate the local independence assumption, after correction for local effects [@qu1996; @schumacher2016].--> For demonstration, we visualised class-wise predicted probability density plots depicting how separable predicted values are between two classes. We also used the diagnosis at discharge (if available) as a pseudo-gold standard to visualise the correlation of our model-based prediction and delayed clinical diagnosis. 

```{r mv-priors, fig.cap="Density plots for prior distributions and their adherence to prior knowledge of sensitvity and specifity for TBM confirmation tests, against then-made clinical diagnosis. Note that Thwaites 2004 was descriptive only while ZN and Culture in Nhu 2013 were references hence no Confidence Interval", warning=FALSE,  fig.align='center', out.width='90%', fig.id="mv-priors", fig.height=8, dpi=300 }
sen_tbl <-
  tibble::tibble(
    'Test' = rep(c('ZN', 'Culture', 'Xpert'), 3),
    'Study' = rep(c('Thwaites 2004', 'Nhu 2013', 'Heemskerk 2018'), each=3),
    'est' = c(58/100, 64/100, NA, 78.64/100, 66.54/100, 59.34/100, 34.54/100, 31.84/100, 25.14/100),
    'lower.ci' = c(NA, NA, NA, 71.94/100, 59.14/100, 51.84/100, 29.94/100, 27.34/100, 21.04/100),
    'upper.ci' = c(NA, NA, NA, 84.34/100, 73.34/100, 66.54/100, 39.44/100, 36.74/100, 29.74/100)
  )

spc_tbl <-
  tibble::tibble(
    'Test' = rep(c('ZN', 'Culture', 'Xpert'), 3),
    # Test_id = rep(c(3,1,2), 2),
    'Study' = rep(c('Thwaites 2004','Nhu 2013', 'Heemskerk 2018'), each=3),
    'est' = c(NA, NA, NA, 0, 0, 0.05/100, 0, 0, 0),
    'upper.ci' = c(NA, NA, NA, NA, NA, (100-97.2)/100, (100-97.1)/100, (100-96.9)/100, (100-96.1)/100),
    'lower.ci' = c(NA, NA, NA, NA, NA, 0, 0, 0, 0)
  )

spc_rng = rbind(
  data.frame(
    Test = 'ZN',
    Test_id = 1,
    logit = rlogis(1000000,qlogis(.001),1.1),
    linear = rlogis(1000000,qlogis(.001),1.1) |> plogis()
  ),
  data.frame(
    Test = 'Culture',
    Test_id = 2,
    logit = rlogis(1000000,qlogis(.001),1.1),
    linear = rlogis(1000000,qlogis(.001),1.1) |> plogis()
  ),
  data.frame(
    Test = 'Xpert',
    Test_id = 3,
    logit = rlogis(1000000,qlogis(.005),.7),
    linear = rlogis(1000000,qlogis(.005),.7) |> plogis()
  )
) |>
  filter(linear<.1)
  

sen_rng = rbind(
  data.frame(
    Test = 'ZN',
    Test_id = 1,
    logit = rlogis(500000, 0,.3),
    linear = rlogis(500000, 0,.3) |> plogis()
  ),
  data.frame(
    Test = 'Culture',
    Test_id = 2,
     logit = rlogis(500000, 0,.3),
    linear = rlogis(500000, 0,.3) |> plogis()
  ),
  data.frame(
    Test = 'Xpert',
    Test_id = 3,
    logit = rlogis(500000, 0,.3),
    linear = rlogis(500000, 0,.3) |> plogis()
  )
) 

spc_linear_plt <- ggplot() + 
  ggdist::stat_halfeye(mapping=aes(y=linear), data=spc_rng) +
  geom_point(aes(y= est, color = Study, x = -.2), data=spc_tbl, position=position_dodge(.2), shape=18, size=3) + 
  geom_linerange(aes(ymin = lower.ci, ymax = upper.ci, color = Study, x = -.2), data=spc_tbl, position=position_dodge(.2)) +
  coord_flip(ylim=c(0,.05))+
  facet_grid(Test~.)+
  xlab('') + ylab('') +
  scale_color_discrete(drop=FALSE)+
  theme_minimal() +
  theme(text = element_text('serif', size = 9), 
        plot.tag = element_text('serif', size = 9),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

spc_logit_plt <- ggplot() + 
  ggdist::stat_halfeye(mapping=aes(y=logit), data=spc_rng) +
  facet_grid(Test~.)+
  xlab('') + ylab('') +
  coord_flip(ylim=c(-15, 2.5)) +
  theme_minimal() +
  theme(text = element_text('serif', size = 9), 
        plot.tag = element_text('serif', size = 9),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

sen_linear_plt <- ggplot() + 
  ggdist::stat_halfeye(mapping=aes(y=linear), data=sen_rng) +
  geom_point(aes(y= est, color = Study, x = -.2), data=sen_tbl, position=position_dodge(.2), shape=18, size=3) + 
  geom_linerange(aes(ymin = lower.ci, ymax = upper.ci, color = Study, x = -.2), data=sen_tbl, position=position_dodge(.2)) +
  coord_flip()+
  facet_grid(Test~.)+
  xlab('') + ylab('') +
  theme_minimal() +
  theme(text = element_text('serif', size = 9), 
        plot.tag = element_text('serif', size = 9),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

sen_logit_plt <- ggplot() + 
  ggdist::stat_halfeye(mapping=aes(y=logit), data=sen_rng) +
  facet_grid(Test~.)+
  xlab('') + ylab('') +
  coord_flip() +
  theme_minimal() +
  theme(text = element_text('serif', size = 9), 
        plot.tag = element_text('serif', size = 9),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
  

plt <- (spc_linear_plt + ggtitle('FPR (1-Specificity)') + theme(legend.position = "none") | spc_logit_plt) /
  (sen_linear_plt + ggtitle('TPR (Sensitivity)') + theme(legend.position = "bottom") | sen_logit_plt) 

for (i in 1:2) plt[[i]] <- plt[[i]] + plot_layout(tag_level = 'new')
color_me <- list("#000000", "#E69F00", "#56B4E9", c("#000000", "#E69F00", "#56B4E9", "#009E73"))
withr::with_options(
  list(ggplot2.discrete.colour = color_me),

  plt + plot_annotation(tag_levels = list(c('',''), 'A'), caption='A: Linear scale, B: Logistic Scale \n Black dots, thick lines and thin lines are median, IQR, and 95% inter-percentile range') + plot_layout(guides = 'collect') & theme(text = element_text('serif', size = 9), plot.tag = element_text('serif', size = 9), legend.position = "bottom") 
)
```

### Sensitivity and exploratory analysis

Since latent class analysis models unobserved characteristics, results may be strongly dependent on our assumptions. Therefore, we conducted sensitivity analyses where some assumptions were lifted or changed. We first increased the standard deviation of our prior for specificities of all confirmatory tests so that every value in the range of 90%-100% were accepted. We also changed the prior distributions of the parameters in the prevalence and bacillary burden models from *normal* to a more skeptical one *Student's t* with 4 degrees of freedom; means and scales, however, were kept as-is. Their *elpd*s were compared. Thirdly, as recent studies suggested a sup-optimal specificity of Xpert test on CSF samples [@nhu2013; @chen2020], we considered a Missing-At-Random scenario, where observation chance of confirmation tests depend on the unknown TBM status and locally independent to the value of confirmation tests. Observation status was then included in the model as a separated manifest variables. We visualised the estimates of this model with the best performing one in our main analysis.

Furthermore, to explore hidden effects, we added quadratic terms for all five CSF bio-markers and RGCS to the prevalence model; CSF volume was additionally included to capture its potential impact on test sensitivity. *Laplace* priors were utilised instead of *Normal* for all linear covariates. Posteriors and performance metrics of this model were reported.

### Simplified classification model

We lastly developed a simplified linear model approximating the previously estimated latent probabilities but without laboratory features. To achieve this, we transferred the analytic forms of posteriors outputted by the best performing model, on the logit scale, to a separate linear regression where the number of predictors were reduced and error terms were assumed to follow standard *logistic* distribution.

All data preparation, cleaning, and processing were performed on statistical package R, version 4.1.1 [@rcoreteam]. The model was developed on the probabilistic language Stan via the interface RStan , version 2.27 [@stan]. Plotting was done using package bayesplot [@bayesplot], classifierplots [@classifierplots], and ggvenn [@ggvenn]. Some other packages used include: R6 [@r6] and rms [@rms]. All estimations came from 8000 effective iterations of 4 MCMC chains. Convergence and effectiveness of MCMC chains were evaluated by Brooks-Gelman-Rubin statistic ($\hat{R}$) and Effective Sample size[@stan-doc]. Mathematical formulations and technical adjustments were detailed in the supplementary documents. All code were published on [project's github repo](github.com/TBM-LCA).

