\clearpage
\pagebreak
\newpage

# (APPENDIX) Supplementary Documents {.unnumbered}

# Statistical supplementary document {.unnumbered}

<!-- :::{custom-style="Title"} -->
<!-- Supplementary document -->
<!-- ::: -->
\pagenumbering{roman}
\setcounter{table}{0}
\setcounter{figure}{0}
\setcounter{page}{0}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

# Model Specification Details {#appendix-model-spec}

<!-- ## Overview -->
## Mathematical parametrisation {#mathematical-parametrisation}

Our model consists of two components:

1.  *Prevalence model*: A logistic regression model for the prevalence of TBM ($\theta$) on a set risk factors ($X$):
  $$
    logit(\theta) \sim X^T\alpha
  $$
  
*where $X$ and $\alpha$ are vectors of covariate values and coefficients.*

2.  *Indicator model*: three logistic regression models for the probability of having a positive test result depending on TBM status. We extended this model as described in Table \@ref(tab:model-archs). Similar to previous studies [@qu1996; @hadgu2002; @schumacher2016], we allowed the test results to depend on the unobserved bacillary burden and systematic noise of the procedures each of which were in turn regressed on relevant bio-markers and an unobserved random effect. Presumably, TBM patients with lower bacillary burden are less likely to be tested positive. Our models 1,2 and 5 are similar in spirit to models *M1*, *M3* and *M4* in [@schumacher2016]. In their models, they didn’t make the direct link between the covariates and bacillary burden, which is what we do in our models 3 and 4. This yields two extra models that stay in-between M3 and M4. 

The structures of our LCMs are formulated in details below. For $1 \leq i \leq 660$, $C \in \{0,1\}$, and $\theta_i = P(C_i=1)$ are the *true TBM status* and the *individual probability of TBM*; $y^{ZN\mbox{-}Smear}_i$, $y^{MGIT}_i$, and $y^{Xpert}_i$} are the *observed test outcomes*. The joint distribution of the observed test results is:

$$
\begin{aligned}
P_i &= P(y^{ZN\mbox{-}Smear}_i,y^{Mgit}_i,y^{Xpert}_i) \\
&= \theta_i * P(y^{ZN\mbox{-}Smear}_i,y^{Mgit}_i,y^{Xpert}_i|C_i=1)\\ 
 &\ \ \ \ + (1-\theta_i) * P(y^{ZN\mbox{-}Smear}_i,y^{Mgit}_i,y^{Xpert}_i|C_i=0) 
\end{aligned} 
$$ {#eq:lca-fml}

Under the local independence assumption that all individuals with the same TBM status have equal probability to test positive (model 1), we obtain:

$$
\begin{aligned}
P_i &= \theta_i * P(y^{ZN\mbox{-}Smear}_i | C=1) * P(y^{Mgit}_i | C=1) * P(y^{Xpert}_i | C=1) \\
&\ \ \ \ + (1-\theta_i) *  P(y^{ZN\mbox{-}Smear}_i | C=0) * P(y^{Mgit}_i | C=0) * P(y^{Xpert}_i | C=0)
\end{aligned}
$$ {#eq:classic-lca-fml}

For models 2 and 3, we assumed that the local independence assumption holds conditionally on TBM status and on bacillary burden ($B$) [@schumacher2016]. Bacillary burden was defined as a standardised, no unit quantification and in turn depends on a set of features (denoted as aggravating factors $V_i^{(t)}$) via a linear regression. [@eq:lca-fml; @eq:classic-lca-fml] become:

$$
\begin{aligned}
P_i &= P(y^{ZN\mbox{-}Smear}_i,y^{Mgit}_i,y^{Xpert}_i) \\
&= \theta_i * P(y^{ZN\mbox{-}Smear}_i,y^{Mgit}_i,y^{Xpert}_i|C_i=1, V^{(t)}_i)\\ 
&\ \ \ \ + (1-\theta_i) * P(y^{ZNvSmear}_i,y^{Mgit}_i,y^{Xpert}_i|C_i=0) \\
&= \theta_i * P(y^{ZN\mbox{-}Smear}_i | C_i=1,V^{(t)}_i) * P(y^{Mgit}_i | C_i=1,V^{(t)}_i) * P(y^{Xpert}_i | C_i=1, V^{(t)}_i) \\
&\ \ \ \ + (1-\theta_i) *  P(y^{ZN\mbox{-}Smear}_i | C_i=0) * P(y^{Mgit}_i | C_i=0) * P(y^{Xpert}_i | C_i=0)
\end{aligned} 
$$ {#eq:re-lca-fml}

We use a mixed-effect logistic regression to capture the correlation between $B$ and test results ($y^{(t)}$):

$$
\begin{aligned}
logit(P(y^{(t)}_i) | C_i = 1, V_i^{(t)}) &= z^{(t)}_0 + B_i * \beta^{(t)} \\
&= z^{(t)}_0 + (V_i^T \gamma + r_i) \beta^{(t)}
\end{aligned} 
$$ {#eq:bd}

where $(t) \in$ \{ZN-Smear, Mgit, Xpert\}, $z^{(t)}_0$ is the intercept, and $\beta^{(t)}_{B}$ is the coefficient of bacillary burden on the result of test (t), $\gamma$ is the vector of coefficients for $V_i$ - in model 2, $\beta^{(t)}$ are assumed to be equal between $t$, and individual random effect $r_i$ quantifies the fluctuation of bacillary burden not explained by $V_i$. $V_i$ and $r_i$ are assumed to be independent of each other. Compared to the previous study [@schumacher2016], this design guarantees that for each pair of features, although their impact on test results could be different, their relative contribution (measured by the pairwise ratios their corresponding parameters) are constrained to be identical:

$$
\frac{\gamma^{[k]} \beta^{(t)}}{\gamma^{[l]} \beta^{(t)}} = \frac{\gamma^{[k]}}{\gamma^{[l]}}
$$

*for $\gamma^{[k]}$ and $\gamma^{[l]}$ are two element of vector $\gamma$ and the coefficient for aggravating factor $V_i^{[k]}$ and $V_i^{[kl]}$.*

In model 4, we additionally captured the sample-wise fluctuation that cannot be explained by the bacillary burden alone (@eq:model-4). This fluctuation was modelled by one extra random effect ($U$) unique for each individual that is not related to the aggravating factors $V$. Its coefficient was represented by the vector $\upsilon$.

$$
\begin{aligned}
logit(P(y^{(t)}_i) | C_i = 1, V_i^{(t)}) &= z^{(t)}_0 + B_i * \beta^{(t)} + U_i^T \upsilon_u^{(t)} \\
&= z^{(t)}_0 + (V_i^T \gamma + r_i) \beta^{(t)} + U_i^T \upsilon_u^{(t)}
\end{aligned} 
$$ {#eq:model-4}

In model 5 (@eq:model-5), we explored all potential impacts of $V$ to test results by linking them directly to $P(y^{(t)}_i) | C_i = 1, V_i^{(t)}$. This is the most flexible design and most similar to model *M4* in [@schumacher2016].

$$
\begin{aligned}
logit(P(y^{(t)}_i) | C_i = 1, V_i^{(t)}) &= z^{(t)}_0 + V_i^T \beta + r_i * b^{(t)}_r 
\end{aligned} 
$$ {#eq:model-5}

Model performance was estimated and compared using expected log pointwise density (elpd) [@vehtari2016]. This measure quantifies how likely a new observation ($y_{new}$) can be observed given the data $y$, model $\mathcal{M}$, and posterior distributions of the parameters $\theta$. A larger elpd means that the predicted posterior distribution concentrates around the observed response value. 

All elpd, ROC curves, AUCs, and calibration metrics are calculated using 5 repetitions of 20-fold cross validation. For each repetition, we pooled all held-out folds together to reconstruct a complete dataset, with their predictions coming from the model trained on the complement of the held-out folds. The averaged performance measures was acquired by pooling all held-out folds over the 5 repetitions together.

```{r model-archs, tab.cap="Different model designs", tab.id="model-archs"}
model_archs <-
  data.frame(
    Model=1:5,
    Def=c(
      'No bacillary burden; everyone in the same TBM class has equal risk to test positive',
      'Test results depend on individual bacillary burden; impacts of bacillary burden on test results is the same across all tests',
      'Impacts of bacillary burden differs between tests',
      'Added technical fluctuation as a second random effect; fixed effects only contributes to bacillary burden',
      'Fixed effects also contribute to technical fluctuation (i.e. direct associtations between V and P(Y))'
    )
  ) 

model_archs |> 
  flextable::flextable() |>
  flextable::set_header_labels(
    Model='Model', 
    Def='Base description (Only additional effects compared to lower number are mentioned)'
  ) |>
  flextable::width(j=1, width=.5)|>
  flextable::width(j=2, width=5) |>
  flextable::theme_vanilla()
```

## Simplified model

The simplified model is a logistic regression that fits a simplified set of risk factors that does not involve a lumbar puncture. The response variable for this simplified model was generated from the posterior distribution of TBM prevalence $\theta$ fitted from the selected model above.

$$
\begin{aligned}
logit(\hat{Y}) &\sim U^T\kappa \\
\hat{Y} &\sim Bernoulli(\theta)
\end{aligned}
$$

*where $\kappa$ is the vector of coefficients for the covariates $U$*

To do this in Stan, we simulated 400 $\hat{Y}$ and combined all MCMC chains into one model. The cross validation procedure for this model was done in the same manner; as we simulated a different response vector $\hat{Y}_i^k$ for every fold $k$ in each repetition $i$ in which $1 \leq k \perp\!\!\!\!\perp i \leq 20$. The final AUC, ROC, and calibration were averaged over these 400 experiments, in which the confidence interval or confidence band was provided.

## Preprocessing and imputation {#appendix-data}

### Data preprocessing

The inclusion of predictors was mainly based on the uniform definition [@marais2010]. Prior knowledges of different predictors were shown in table \@ref(tab:predictor-tab). Days from onset to admission and CSF biomarkers were transformed to logarithmic scale as they were count data and right-skewed. We followed Gelman's scale-matching method [@gelman2008], in which continuous predictors were centred by their empirical means and subsequently scaled by 2 times their standard deviations (sd) while binary ones were kept as-is. CSF eosinophil count was an exception, as it was heavily zero-inflated, we created an extra indicator for count > 0 and rescaled the positive values by their sd. Glasgow Coma Score (GCS) and its scale components (Voice - GCSV, Eyes - GCSE, and Muscle - GCSM) were translated to *Reversed GCS* (RGCS = RGCSV + RGCSE + RGCSM) so that a $GCS = 15$ would be equivalent to $RGCS = 0$, while $GCS = 3$ would be translated to $RGCS = 12$ (table \@ref(tab:gcs-tab)). Binary variables were dummy-coded into 0 for "Negative" / "No" and 1 for "Positive" / "Yes". To facilitate the imputation model, we scaled down the auxiliary variables RGCSE, RGCSM, and RGCSV down to a full range of $[0, 1]$. Hence $RGCS = 3 RGCSE + 4 RGCSV + 5 RGCSM$.

```{r gcs-tab, out.width='100%', tab.id='gcs-tab', tab.cap='Conversion table from classic Glasgow Coma Scale to Reversed GCS (RGCS)'}

tibble::tribble(
  ~ Feature,                     ~ Response, ~ GCS, ~ RGCS,
  'Eye response',        'Open sponatenously', 4, 0,
  'Eye response',     'Open to voice command', 3, 1,
  'Eye response',              'Open to pain', 2, 2,
  'Eye response',               'No eye open', 1, 3,
  'Verbal response',             'Orientated', 5, 0,
  'Verbal response',               'Confused', 4, 1,
  'Verbal response',     'Inappopriate words', 3, 2,
  'Verbal response','Incomprehensible sounds', 2, 3,
  'Verbal response',     'No verbal response', 1, 4,
  'Motor response',            'Obey command', 6, 0,
  'Motor response',         'Localising pain', 5, 1,
  'Motor response',    'Withdrawal from pain', 4, 2,
  'Motor response',                 'Flexing', 3, 3,
  'Motor response',               'Extending', 2, 4,
  'Motor response',       'No motor response', 1, 5
) |>
  flextable::flextable() |>
  flextable::merge_v(j = 1) |>  
  flextable::width(j=1, width=2) |>
  flextable::width(j=2, width=2) |> 
  flextable::theme_vanilla()
```

### Imputation strategy {#stat-impute}

Our imputation models are summarised in Figure \@ref(fig:impute-model), together with their auxiliary variables. In general, for each incomplete variable, the missing values were sampled from a distribution fitted to corresponding observed values, as part of the MCMC sampling. Composite variables (*TB-suggested symptoms*, *Local neurological deficit*, and *GCS*) were imputed per their corresponding compartments. Potentially correlated variables were grouped together and sampled from a multivariate distribution. Due to Stan not supporting Multivariate Logistic Distribution, all binary predictors were imputed using Probit models [@stan-doc, @albert1993] for consistency. Numeric predictors were imputed using Multivariate Linear Regression. Note that due to technical limitation, imputed RGCS were treated as if they were continuous in the sampling step, but were rounded to nearest integers when estimating the model performance. Hence, we assumed:

$$
\begin{aligned}
 X_{cont} &\sim \mathcal{N}(U_{X_{cont}}^TG, \Sigma) \\
 Probit(X_{binary}) &\sim \mathcal{N}(U_{X_{binary}}^TG, I)
\end{aligned}
$$

*where $X_{cont}$ and $X_{binary}$ are vectors of continuous and binary variables to be imputed, and $U^T_{X_{cont}}$ and $U^T_{X_{binary}}$ are vector of auxiliary covariates used in the respective imputation model, with coefficients $G$. $\mathcal{N}$ is the *Normal *distribution in univariate case and *Multivariate Normal *distribution otherwise with $\Sigma$ as the variance-covariance matrix; in the binary case, $\Sigma$ is the identity matrix $I$ [@albert1993].*

```{dot impute-model, echo=FALSE, out.width="100%", fig.align="center", fig.cap="Imputation strategy for predictors. Variables in white boxes with solid borders were used in the model. Variables in ovals with dashed borders were only contributed in the imputation model as donors and were not directly included in the main model. Variables in the same grey boxes were imputed together via a multivariate regression. Arrows demonstrate a covariate $\\rightarrow$ response relation. \"Mycobacterial test done\" is defined as ZN-Smear, MGIT and Xpert were done for the patient.", fig.id="impute-model"}

digraph imp_model{
  rankdir = TB;
  resolution=500;
  compound=true;
  # concentrate=true;
  splines=ortho;
  graph[fontname="CMU-Serif"];
  edge[fontname = "CMU-Serif"];
  node[fontname = "CMU-Serif"];
  
  subgraph cluster_csf{
    rank=same;
    bgcolor = lightgrey;
    node[style=filled fillcolor=white];
    subgraph glu{
      peripheries=0;
      CsfGlu;
      CsfLymp;
    }
    BldGlu;
    CsfNeu;
    subgraph bio{
      peripheries=0;
      CsfPro;
      CsfLac;
    }
  }
  
  subgraph cluster_clinsymp{
    rank=same;
    label = "TB-suggested symptoms";
    bgcolor = lightgrey;
    node[style="filled,dashed" fillcolor=white];
    Cough[label="Coughing"];
    subgraph lowerz{
      peripheries=0;
      Nsweats[label="Night Sweats"];
      WLoss[label="Weight Loss"];
    }
  }
  
  subgraph cluster_motor{
    rank=same;
    label = "Local motor deficit";
    bgcolor = lightgrey;
    node[style="filled,dashed" fillcolor=white];
    Hemi[label="Hemiplegia"];
    subgraph lowerz{
      peripheries=0;
      Para[label="Paraplegia"];
      Tetra[label="Tetraplegia"];
    }
  }
  
  subgraph cluster_gcs{
    label = "Reversed Glagow Coma Score";
    bgcolor = lightgrey;
    node[style="filled,dashed" fillcolor=white];
    RGCSV; RGCSM; RGCSE;
  }
  # 
  # subgraph cluster_bld{
  #   # peripheries=0;
  #   bgcolor = lightgrey;
  #   rank=same;
  #   node[style="filled,dashed" fillcolor=white];
  # }
 
  obs[label="Mycobacterial test done", shape="box"];
  HIV[label="HIV Status", shape="box"];
  # Age[label="Age", shape="box"];
  TBDays[label="Illness days", shape="box"]; 
  BldGlu[label="Blood Glucose", shape="box"];
  CsfGlu[label="CSF Glucose", shape="box"]; 
  CsfLymp[label="CSF Lymphocytes", shape="box"];
  CsfNeu[label="CSF Neutrophils", shape="box"];
  CsfPro[label="CSF Protein", shape="box"];
  CsfLac[label="CSF Lactate", shape="box"];
  # GCS[label="Glasgow Coma Score", shape="box"];
  
  obs -> HIV
  obs -> TBDays;
  obs -> Hemi[lhead=cluster_motor];
  obs -> Cough[lhead=cluster_clinsymp];
  obs -> RGCSV[lhead=cluster_gcs];
  obs -> CsfGlu[lhead=cluster_csf];
  HIV -> TBDays;
  HIV -> Hemi[lhead=cluster_motor];
  HIV:w -> Cough[lhead=cluster_clinsymp];
  HIV -> RGCSV[lhead=cluster_gcs];
  HIV -> CsfLymp[lhead=cluster_csf];
  
  TBDays -> Cough[lhead=cluster_clinsymp];
  TBDays -> Hemi[lhead=cluster_motor];
  
  Cough -> Nsweats[dir=none];
  Nsweats -> WLoss[dir=none];
  WLoss -> Cough[dir=none];
  Hemi -> Para[dir=none];
  Para -> Tetra[dir=none]; 
  Tetra -> Hemi[dir=none];
  
  CsfGlu -> BldGlu[dir=none];
  CsfGlu -> CsfLymp[dir=none];
  CsfLymp -> CsfNeu[dir=none];
  CsfNeu ->CsfPro[dir=none];
  CsfPro -> CsfLac[dir=none];
  BldGlu -> CsfLac[dir=none];

  RGCSV -> RGCSE[dir=none];
  RGCSE -> RGCSM[dir=none];
  RGCSM -> RGCSV[dir=none];
}  
```

## Prior choices {#appendix-stat-prior-choices}

### Main model

For the prevalence model, we chose Student's *t* distribution with 6 degrees of freedom ($t_6$) for intercept, in which we fixed the mean to 0 and scale to 3 because its absolute value was unlikely to be higher than 6 (which is equivalent to an probability range of (0.002 - 0.998)) [@boonstra2019]. For the regression coefficients, we used the same *t* distribution family with means set to 0. For the scales we incorporated a penalty term sampled from $Normal(0, 1.2)$ as we did not expect coefficient norms to be larger than 4.8 in absolute value (which is equivalent to odd ratio range of 0.008 - 121.510) [@prior-choice, @vanerp2019], but we left some room for our expectations to be wrong. For high-impact risk factors (marked as "strong" in Table \@ref(tab:predictor-tab)), priors were further truncated at 0, so that only values in the expected direction were accepted in the MCMC chains. We rescaled continuous variables by their post-imputed sd. 
$$
\begin{aligned}
s &\sim Normal(0, 1.2) \\
a_0 &\sim t_6(0, 3)\\
a_{binary} &\sim t_6(0, s)\\
a_{continuous} &\sim t_6(0, \frac{s}{2sd})
\end{aligned}
$$

*where $s$ is the adaptive penalty term, $a_0$, $a_{binary}$, and $a_{continuous}$ are the intercept, coefficients for the binary covariates , and continuous covariates included in the prevalence model. $sd$ is the empirical standard deviation of each continuous variables after the imputation and is different between those variables.* 

$$
\begin{aligned}
  1-Spc_{Xpert} &\sim Logistic(logit(.005), .7) \\
  1-Spc_{MGIT}  &\sim Logistic(logit(.001), 1) \\
  1-Spc_{ZN\mbox{-}Smear} &\sim Logistic(logit(.001), 1) \\ 
  Sen_{Xpert,\ MGIT,\ ZN\mbox{-}Smear} &\sim Logistic(0,.35)
\end{aligned}
$$ {#eq:priors_response}

For the indicator model, our choice of priors was based on information collected from several previous studies [@nhu2013, @thwaites2004, @heemskerk2018]. A summary of these choices is shown in Figure \@ref(fig:mv-priors)), on logistic and linear scale, and how good they covered corresponding results derived from the previous studies. As suggested by the literature, we used highly informative priors for False Positive Rate ($FPR = 1-Specificity$) and weakly informative priors for True Positive Rate ($TPC = Sensitivity$) on the logit scale (@eq:priors_response), given the discrepancies between the studies.

We generally applied the same rules used in the prevalence model to choose prior distribution for coefficients $\gamma$ that take part in latent bacillary burden model $(B_i = V_i^T\gamma* + r) * \beta^{(t)} + u_i * \beta_{u}^{(t)}$ where $u$ only exists in Model 4. Both latent random effect $r$ and $u$ were assumed following *Standard Normal* distribution. To avoid sign switching, we constrained $\beta^{(t)}$ to be non-negative As coefficients $\gamma$ and $\beta^{(t)}$ were multiplied together, to provide a fair penalisation, we divided the penalty term $s_2$ for $\gamma$ by the average of $\beta^{(t)}$ over three tests *t = {ZN-Smear, Mgit, Xpert}*.

$$
\begin{aligned}
r &\sim Normal(0, 1) \\
u &\sim Normal(0, 1) \\
s_2 &\sim Normal(0, 1.2) \\
\beta^{(t)} &\sim Normal(0, s_2) \\
\gamma_{HIV} &\sim t_5(0, s_2) \\
\gamma_{continuous} &\sim t_5(0, \frac{s_2}{sd * \overline{\beta^{(t)}}})
\end{aligned}
$$

*where $\gamma_{HIV}$ is the coefficient for HIV status, $\gamma_{continuous}$ is the vector of coefficients for other CSF biomarkers, $s_2$ is the penalty term for the bacillary burden model. $\overline{\beta^{(t)}}$ is the average of $\beta^{(t)}$ over (t). $sd$ is the empirical standard deviation of each continuous variables after the imputation and is different between those variables.*

### Imputation model

For imputation, $Normal(0, 2.5)$ was chosen as prior distributions for intercepts and coefficients [@prior-choice], except for RGCS (E, V, and M) as their domains were two-sided constrained (section \@ref(data-preprocessing)) - in which case, the mode of RGCSV, RGCSM, and RGCSE were sampled from an $Exponential(\beta = 20)$ truncated within [0, 1], and $Normal(0, 0.25)$ was chosen as the prior distribution for their sd.

In cases where imputed variables were assumed multivariably correlated, Cholesky decomposition of the variance-covariance matrix was sampled from a *Cholesky Lewandowski-Kurowicka-Joe (LKJ) Correlation* prior with $\eta = 4$ [@stan-doc].

# Model diagnosis and comparision

## Estimated log-pointwise density of each considered model

```{r elpd-tbl, tab.id='elpd-tbl', tab.cap = 'Expected elpd of considered models and their standard errors'}
elpd =  readRDS(file.path(data_dir, '..', 'export/metrics/elpd.RDS'))
elpd_tbl =
  data.frame(
    Model = 1:5,
    elpd = sapply(elpd, \(x) x$estimates[,'Estimate'] |> formatC(format = 'f', digits=2)),
    SE = sapply(elpd, \(x) x$estimates[,'SE']  |> formatC(format = 'f', digits=2))
  )

elpd_tbl |>
  flextable::flextable()|>
  flextable::width(width=1.5) |>
  flextable::set_header_labels(SE = 'Standard error') |>
  flextable::highlight(i = 3, j = 2) |>
  flextable::theme_vanilla()
```

## NUT Sampler diagnosis for the selected model

The histogram of Brook-Gelman-Rubin $\hat{R}$ statistic, effective sample size, and Monte Carlo standard errors across all parameters are shown in \@ref(fig:nut-dx) [@stan-doc]. Overall, no sign of ill samples were suggested by the diagnostics plots, with most $\hat{R}$ highly concentrated around $1 \pm .005$; with effective sample size $\geq$ 50% of the whole data.

```{r load_data, warning=FALSE, include=FALSE}
stanDx <- readRDS(file.path(data_dir, '../export/metrics/stanDx.RDS'))
```

```{r nut-dx, fig.align='center', fig.cap="NUT Sampler diagnostic statistics for the selected model", fig.id="resid-pwcorr", warning=FALSE, out.width='60%', fig.width=6, fig.asp=2, message=FALSE}

wrap_plots(stanDx, ncol=1)
```


# Assumption check and sensitivity analysis {#appendix-sen-analysis}

## Methods

### Latent class model 

Apart from model performance metrics, we also checked for our conditional independence assumptions, without (model 1) and with correction for latent bacillary burden (model 2+). To do this, we plotted the residual pairwise correlation between confirmatory tests, effectively comparing predicted pairwise correlation and observed pairwise correlation [@qu1996].

As observed by a previous study [@nhu2013], GeneXpert might not have perfect specificity, hence our assumption of imputing missing Xpert as Negative might not hold. To check this, we did a sensitivity analysis where we incorporated the incompleteness of Xpert into our model. We assumed observation (or missingness) of Xpert was dependent and only dependent on TBM allocation of the patients. We hence included it as a fourth manifest variable, denoted as *obs*, apart from ZN-Smear, MGIT, and Xpert. When Xpert is missing, $obs = 0$, and vice versa. The likelihood function (@eq:lca-fml) hence becomes:

$$
\begin{aligned}
P_i &= P(y^{ZN\mbox{-}Smear}_i,y^{Mgit}_i,y^{Xpert}_i, {obs}_i)\\
&= \theta_i * P(obs_i) * P(y^{ZN\mbox{-}Smear}_i,y^{Mgit}_i,y^{Xpert}_i|C=1, obs=1)\\ 
&\ \ \ \ + \theta_i * (1-P(obs_i)) * P(y^{ZN\mbox{-}Smear}_i,y^{Mgit}_i|C=1, obs=0)\\ 
&\ \ \ \ + (1-\theta_i) * P(obs_i) * P(y^{ZN\mbox{-}Smear}_i,y^{Mgit}_i,y^{Xpert}_i|C=0, obs=1) \\
&\ \ \ \ + (1-\theta_i) * (1-P(obs_i)) * P(y^{ZN\mbox{-}Smear}_i,y^{Mgit}_i|C=0, obs=0) 
\end{aligned} 
$$ {#eq:lca-missingXpert-fml}

given that _obs_ is locally independent to the three confirmatory tests. We used a weakly informative prior distribution for non-TBM group and an informative prior distribution for TBM group (Supplementary Figure \@ref(fig:prior-obs)) as we believed most patients in the TBM group have been tested with confirmatory tests. However, we widen the scale in the latter case enough so that an extreme value of 20% TBM patients who left untested were allowed.

```{r prior-obs, fig.align='center', out.width='100%', warning=FALSE, message=FALSE, fig.cap = "Prior distribution for Xpert response rate, in non-TBM patients (left) and TBM patients (right). Point, thick lines, and thin lines are the median, 50\\% and 95\\% inter-quantile ranges."}

obs_rng = rbind(
  data.frame(
    Test = 'Xpert observed / non-TBM',
    Test_id = 1,
    logit = rlogis(500000, 0,.5),
    linear = rlogis(500000, 0,.5) |> plogis()
  ),
  data.frame(
    Test = 'Xpert observed / TBM',
    Test_id = 2,
    logit = rlogis(500000, qlogis(.99),1),
    linear = rlogis(500000, qlogis(.99),1) |> plogis()
  )
) 


obs_logit_plt <- ggplot() + 
  ggdist::stat_halfeye(mapping=aes(y=logit), data=obs_rng) +
  facet_grid(Test~.)+
  scale_y_continuous(breaks = qlogis(c(0.01, .1, .5, .9, .99, .999)), 
                     labels = c(0.01, .1,.5, .9, .99, .999),
                     limits = qlogis(c(0.0005, .999995))) +
  xlab('') + ylab('') +
  coord_flip() +
  theme_minimal() +
  theme(#text = element_text('serif', size = 9)
        plot.tag = element_text('serif', size = 9),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
        )
obs_logit_plt
  
```
<!--
To look at the validity of this method, we performed a simulation study. In this study, we created $X, X_2, X_3$ as three covariates for latent class $C$. For simplicity, $Y_1, Y_2$, and $Y_3$ were three manifest variables whose data generation processes are locally independent w.r.t. $C$. $Y_3$, however were not completely observed; its observation rate (represented by the binary variable obs_rate) are different for $C=0$ and $C=1$. 

The data generation process was done in **R** version 4.1 [@rcoreteam] as below. We performed 1000 simulations for each scenarios. For each scenario we extracted the expected values from posterior distributions, visualised them to see whether they concentrated around the real values.All models were fitted under the probabilistic language **Stan** version 2.27 [@stan-doc].

``` r
# Misc function
generate_Y = \(C, probs){
  Cs = sort(unique(C))
  sapply(C, \(c) rbinom(1,1, probs[Cs==c]))
}

# Sample size
N        = 1000 

# Create data with 3 predictors, X, X2, and X3
X        =  rnorm(N, 0, 1 )
X2       =  rnorm(N, 0, 1 )
X3       = rbinom(N, 1, .3)

# Create latent class
probs    = plogis(3*X+X2+5*X3-1)
C        = sapply(probs, \(p) rbinom(1,1,p))

# Manifest variables
Y1       = generate_Y(C, c(.1 ,.3))
Y2       = generate_Y(C, c(.01,.5))
Y3       = generate_Y(C, c(.05,.8))

# Simulate class-aware missing data for Y3. If obs3==1, Y3 is observed
obs_rate = c(.1, .95)
obs3     = generate_Y(C, obs_rate)
```

The likelihood function of the model shall be
$$
P(Y_1,Y_2,Y_3,obs_3|X,X_2,X_3) = \prod_{n=1}^N \sum_{c=0}^1 (P_n(Y_1,Y_2,Y_3,obs_3|C=1)P(C=c|X,X_2,X_3)
$$ 
Consider an individual $n$: If $obs3=1$: 
$$
\begin{aligned}
P(y_1=Y_1^{(n)},y_2=Y_2^{(n)},y_3=Y_3^{(n)},obs_3=1|X_n,X_2^{(n)},X_3^{(n)}) &= \sum_{c=0}^1 \prod_{i=1}^3 P(y_i=Y_i^{(n)}|obs_3=1,C=c)P(obs_3=1)P(C=c)\\
&= \sum_{c=0}^1 \prod_{i=1}^3 P(y_i=Y_i^{(n)}|C=c)P(obs_3=1)P(C=c)
\end{aligned}
$$
(as $Y_1, Y_2, Y_3, obs_3$ are conditionally independent on $C$).

Similarly, for those whose $obs_3=0$: $$
\begin{aligned}
P(y_1=Y_1^{(n)},y_2=Y_2^{(n)},obs_3=0|X^{(n)},X_2^{(n)},X_3^{(n)} &= \sum_{c=0}^1 \prod_{i=1}^2 P(y_i=Y_i^{(n)}|obs_3=0,C=c)P(obs_3=0)P(C=c)\\
&= \sum_{c=0}^1 \prod_{i=1}^2 P(y_i=Y_i^{(n)}|C=c)P(obs_3=0)P(C=c)
\end{aligned}
$$

We considered five sets of prior distribution for $X, X_2, X_3, Y_1, Y_2, Y_3$, and $obs_3$ corresponding for the level of prior knowledge. The formulation of priors follow the syntax in (@eq:priors_response)`.

1.  Weak prior:

$$
\begin{aligned}
(1-Spc_{X,X_2,X_3}) &\sim Logistic(0,.7) \\
Sen_{X,X_2,X_3} &\sim Logistic(0,.7) \\
obs\_rate|{C=0} &\sim Logistic(0,.7) \\
obs\_rate|{C=1} &\sim Logistic(0,.7)
\end{aligned}
$$

2.  Good but conservative knowledge for observation rate, weak prior for manifest variables:

$$
\begin{aligned}
(1-Spc_{X,X_2,X_3}) &\sim Logistic(0,.7) \\
Sen_{X,X_2,X_3} &\sim Logistic(0,.7) \\
obs\_rate|{C=0} &\sim Logistic(logit(.1),.7) \\
obs\_rate|{C=1} &\sim Logistic(0,.7)
\end{aligned}
$$

3.  Good but conservative knowledge for observation rate and manifest variables:\
$$
\begin{aligned}
(1-Spc_{X}) &\sim Logistic(logit(.1),.7) \\
(1-Spc_{X_2}) &\sim Logistic(logit(.01),.7) \\
(1-Spc_{X_3}) &\sim Logistic(logit(.05),.7) \\
Sen_{X,X_2,X_3} &\sim Logistic(0,.7) \\
obs\_rate|{C=0} &\sim Logistic(logit(.1),.7) \\
obs\_rate|{C=1} &\sim Logistic(0,.7)
\end{aligned}
$$

4.  Very strong knowledge for observation rate, weak prior for manifest variables:\
$$
\begin{aligned}
(1-Spc_{X,X_2,X_3}) &\sim Logistic(0,.7) \\
Sen_{X,X_2,X_3} &\sim Logistic(0,.7) \\
obs\_rate|{C=0} &\sim Logistic(logit(.1),.25) \\
obs\_rate|{C=1} &\sim Logistic(0,.7)
\end{aligned}
$$

5.  Very strong knowledge for observation rate and manifest variables:\
$$
\begin{aligned}
(1-Spc_{X}) &\sim Logistic(logit(.1),.2) \\
(1-Spc_{X_2}) &\sim Logistic(logit(.01),.2) \\
(1-Spc_{X_3}) &\sim Logistic(logit(.05),.2) \\
Sen_{X,X_2,X_3} &\sim Logistic(0,.7) \\
obs\_rate|{C=0} &\sim Logistic(logit(.1),.25) \\
obs\_rate|{C=1} &\sim Logistic(0,.7)
\end{aligned}
$$



```{stan_code simulation-stan-code, pyg.ext='stan'}
data {
  int<lower=1> N;
  int<lower=0> nX;
  matrix[N,nX] X;
  int<lower=0, upper=1> Y1[N];
  int<lower=0, upper=1> Y2[N];
  int<lower=0, upper=1> Y3[N];
  int<lower=0, upper=1> obs3[N];
  int<lower=0, upper=2> good_prior[2]; 
  //0: weak, 1:moderate, 2: strong. First position is for obs_rate, second is for tests
}

parameters {
  ordered[2] z1;
  ordered[2] z2;
  ordered[2] z3;
  ordered[2] zo;
  vector[nX] a;
  real a0;
}

model{
  
  if (good_prior[1]==0){
    zo[1] ~ logistic(0, .7);
    zo[2] ~ logistic(0, .7);
  } else if (good_prior[1]==1) {
    zo[1] ~ logistic(logit(0.01), .7);
    zo[2] ~ logistic(logit(0.95), .7);
  } else {
    zo[1] ~ logistic(logit(0.01), .25);
    zo[2] ~ logistic(logit(0.95), .25);
  }
  
  if (good_prior[2]==0){
    z1    ~ logistic(0,.7);
    z2    ~ logistic(0,.7);
    z3    ~ logistic(0,.7);
  } else if (good_prior[2]==1){
    z1[1] ~ logistic(-2.19,.7);
    z2[1] ~ logistic(-4.59,.7);
    z3[1] ~ logistic(-2.94,.7);
    z1[2] ~ logistic(0,.7);
    z2[2] ~ logistic(0,.7);
    z3[2] ~ logistic(0,.7);
  } else {
    z1[1] ~ logistic(-2.19,.2);
    z2[1] ~ logistic(-4.59,.2);
    z3[1] ~ logistic(-2.94,.2);
    z1[2] ~ logistic(0,.7);
    z2[2] ~ logistic(0,.7);
    z3[2] ~ logistic(0,.7);
  }
  for (n in 1:N){
      real z = a0 + X[n,:]*a;
      real p = inv_logit(z);
      real ll3[2] = 
        obs3[n] == 1 ? 
        {bernoulli_logit_lpmf(Y3[n]|z3[1]), bernoulli_logit_lpmf(Y3[n]|z3[2])} :
        {0, 0};
      target += log_mix(p,
        bernoulli_logit_lpmf(obs3[n]|zo[2]) + 
        bernoulli_logit_lpmf(Y1[n]|z1[2]) + bernoulli_logit_lpmf(Y2[n]|z2[2])+ll3[2],
        bernoulli_logit_lpmf(obs3[n]|zo[1]) + 
        bernoulli_logit_lpmf(Y1[n]|z1[1]) + bernoulli_logit_lpmf(Y2[n]|z2[1])+ll3[1]);
  }
}
```
-->

### Imputation model

To diagnose the validity and impact of our imputation model, we (i) visually compared the distributions of the imputed data and observed data, conditional on the response propensity: to do this, for each variable of interest $Y_v$ first we fitted a logistic regression with the indicator of missingness as the response variable and the completed variables apart from $Y_v$ (denoted as $Y_{-v}$) as the independent variables, then we plot the propensity of response ($e_v$) against $Y_v$, on different colours depending on whether they are imputed values or not; (ii) compared the residual kernel density of $Y_v^{imputed}$ and $Y_v^{observed}$ conditionally on $e_v$; and (iii) performed a posterior predictive check in which we compared the expected coefficients of our completed variables (which contained observed and imputed values) and 100 replicated, fully-imputed variables (which contained only generated values from our imputation model) -- for a good imputation we would expect the likelihood to witness the former is not extreme. The details of the methods are outlined and discussed elsewhere [@nguyen2017; @bondarenko2016]. For simplicity we only analysed features that had more than 10 missing observations.

## Results

### Latent class model

The expected residual correlations between 3 pairs of confirmatory tests for 5 model choices are shown in Supplementary Figure \@ref(fig:resid-pwcorr). Except for model 1 which underestimated the observed values, other models that incorporated random effects showed a good correspondence between predicted and observed correlation, in which case all residual values fluctuated around 0. This suggested that a correction for local dependence was indeed necessary.

```{r resid-pwcorr, fig.align='center', fig.cap="Residual correlation plots for five different model choices", fig.id="resid-pwcorr", warning=FALSE, out.width='50%', fig.width=4, message=FALSE}
rpc_plot = readRDS(file.path(data_dir, '..', 'export', 'metrics', 'rpc_plot.RDS'))
rpc_plot + scale_color_brewer(type='qual', palette = 2) +  theme(legend.position = 'bottom')
```

```{r plot-missing, out.width='100%', fig.dim=c(10, 12.5), fig.align='center', fig.cap='Estimates of model 3 (denoted as Original model) and model 3 with missing Xpert (denoted as Missing model), for prevalence model, bacillary burden model, and sensitivity (TPR) and (1 - Specificity) (FPR) for 3 confimatory tests and the response rates of Xpert for TBM and non-TBM patients. In all plots: dots, thick lines, and thin lines are the means, 50\\%, and 95\\% credible intervals.'}

plot_m <- readRDS(file.path(data_dir, '..', 'export', 'm3m_plot.RDS'))
z_plot <- readRDS(file.path(data_dir, '..', 'export', 'z_plot_u.RDS'))

wrap_plots(plot_m[names(plot_m)!='z_plot'], ncol=1) + z_plot + 
  plot_layout(guide='collect', heights = c(7,3,3)) +
  plot_annotation(tag_levels = 'A') & theme(legend.position='bottom')

```

There were no large difference in the posterior estimates of both models with missing Xpert and the selected model (model 3) (Supplementary Figure \@ref(fig:plot-missing)). The response rate for Xpert in non-TBM group is estimated as `r plot_m$z_plot$data |> filter(parameter == 'z_observe[1]' & Model == "Incomplete Xpert") |> select(m) |> unlist() |> plogis() |> round(2)` (`r plot_m$z_plot$data |> filter(parameter == 'z_observe[1]' & Model == "Incomplete Xpert") |> select(ll) |> unlist() |> plogis() |> round(2)` - `r plot_m$z_plot$data |> filter(parameter == 'z_observe[1]' & Model == "Incomplete Xpert") |> select(hh) |> unlist() |> plogis() |> round(2)`).  

### Imputation of covariates

```{r load_impute_data}
library(ggfx)
load(file.path(data_dir, '..', 'export', 'impute_cv_plot_hex.Rdata'))
```

Residual kernel density against response propensity scores for imputed and observed values in general shows good correspondence (Supplementary Figure \@ref(fig:impute-plot)). We could see there was no obvious difference in imputed and observed values conditioned on response propensity. We also witnessed similar patterns in the residual kernel densities of observed and missing values. The posterior density of feature coefficients and intercept estimated from our completed data corresponded well with those from fully-imputed data. All three figure suggested that our method of imputation were robust and did not severely bias the results.

```{r impute-plot, out.width='75%', fig.dim=c(8,12), fig.align='center', fig.cap = "Diagnosis plot for imputation model. On the left are hexagonal heatmaps between the reponse propensity and the values predicted by the model, stratified by their missingness. On the right hand side are the kernel density plots of the residuals of the linear regression model $y \\sim e_y$ where $y$ is the variable of interest and $e_y$ is the response propensity of that variable."}

hiv_plot$name = 'HIV'
cs_plot$name = 'TB symptoms'
mp_plot$name = 'Focal neurological deficit'
gcs_plot$name = 'GCS'
id_plot$name = 'Duration since onset'
impute_plots <- 
  lapply(
    list(hiv_plot, cs_plot, mp_plot, id_plot, gcs_plot),
    function(plt){
     # wrap_elements(grid::textGrob(plt$name, rot=90)) + 
        wrap_ggplot_grob(plt$plot1)  + wrap_ggplot_grob(plt$plot2) + plot_layout(widths=c(1,1))
      # wrap_elements(gridExtra::arrangeGrob(, nrow=1, left=plt$name))
    }
  )

plt <- wrap_plots(impute_plots, ncol=1) + plot_annotation(tag_levels = list("HIV", "TB_symptoms", "Local_neuro-defict", "Duration since onset", "GCS")) + plot_layout(guides='collect')
  
plt
```

```{r impute-ppc, out.width='100%', fig.dim=c(12,9), fig.cap = "Posterior distributions of features with missing values in the prevalence model, in which the posterior distribution of the selected model (red line) was compared against those of 100 fully-imputed data generated from the fitted imputation model (grey lines)"}
ppc <- readRDS(file.path(data_dir,'../export/ppc_checkresults.RDS'))
vars <- c('HIV' = 'a1', 
          'TB-suggested syndrome' = 'a2',
          'Focal neurological deficit' = 'a3',
          'Days from onset' = 'a11', 
          'GCS' = 'a18',
          'Intercept' = 'a0')
for (a in seq_along(vars)) ppc[[vars[[a]]]] <- patchwork::wrap_elements(ppc[[vars[[a]]]]) + ggtitle(names(vars)[[a]])
patchwork::wrap_plots(ppc, ncol=2, nrow=3) 


```
