---
title: 'In search of a novel scoring system for Tuberculous Meningitis Diagnosis: a Bayesian Latent Class Analysis Approach'
site: bookdown::bookdown_site
bibliography: includes/references.bib
csl: includes/clinical-infectious-diseases.csl
output:
  bookdown::pdf_book:
    base_format: rmarkdown::latex_document
    toc: no
    pandoc_args:
      - '--lua-filter=includes/scholarly-metadata.lua'
      - '--lua-filter=includes/author-info-blocks.lua'
      - '--standalone'
    latex_engine: xelatex
    highlight: pygments
    includes:
      in_header: includes/header.tex
---

```{r setup, include=FALSE}
options(tinytex.engine_args = '-shell-escape')
knitr::opts_chunk$set(echo = FALSE, dpi=300) #, dev = 'png'
flextable::set_flextable_defaults(fonts_ignore=TRUE, font.size=9, table.layout='fixed')
```

```{r library, include=FALSE}
library(gtsummary)
library(dplyr)
library(data.table)
library(magrittr, include.only = '%$%')
library(ggplot2)
library(patchwork)
library(officedown)

# All the misc functions go here.
my_fn <- new.env()

## Add labels accross a tbl
my_fn$add_labels <- \(.data, ...){
  cols <- list(...)
  for (i in seq_along(cols))
    attr(.data[[names(cols)[i]]], 'label') <- cols[[i]]
  .data
}

## Register a knit engine for stan code highlight
my_fn$stan_knitr_engine <- \(options){
  out <- c('\\begin{Shaded}','\\begin{minted}{stan}', options$code, '\\end{minted}','\\end{Shaded}')
  options$echo <- FALSE
  options$results <- "asis"
  knitr::engine_output(options, options$code, out)
}
knitr::knit_engines$set(stan_code = my_fn$stan_knitr_engine)

## Gist to number equations
## https://gist.github.com/mcanouil/eb75057432ff77846f4273d8808e615a
my_fn$is_docx_output <- function (fmt = knitr:::pandoc_to()) {
  if (length(fmt) == 0) {
    return(FALSE)
  } else {
    return(fmt == "docx")
  }
}

my_fn$numberEq.docx <- function (eq, lab, envir = docx.eqcounter) {
  assign(x = "counter", value = get(x = "counter", envir = envir)+1, envir = envir)
  assign(x = gsub("eq:", "", lab), value = get(x = "counter", envir = envir), envir = envir)
  lab <- get("counter", envir = envir)
  return(c('$$', eq, '\\;\\;\\;\\;(', lab, ')', '$$'))
}
my_fn$labEq.docx <- function (lab, envir = docx.eqcounter) {
  return(paste0('(', get(x = gsub("eq:", "", lab), envir = envir), ')'))
}

docx.eqcounter <- new.env()
docx.eqcounter$counter <- 0
  
# Register a knit engine for formula
my_fn$aligned_formula_knit <- \(options){
  out <- my_fn$numberEq.docx(options$code, lab = options$fml.lab)
  options$echo <- FALSE
  options$results <- "asis"
  knitr::engine_output(options, options$code, out)
}
knitr::knit_engines$set(aligned_formula = my_fn$aligned_formula_knit)

# Register diagrammer graphviz
my_fn$diagrammer_engine <- \(options){
  grviz_opt = formals(DiagrammeR::grViz)
  grviz_opt = grviz_opt[names(grviz_opt) != 'diagram']
  new_opt = options[names(options) %in% names(grviz_opt)]
  # not_grviz_opt = opt[!names(opt %in% c('diagram', name(grviz_opt)))]
  grviz_opt = modifyList(grviz_opt, new_opt)
  dia_call <- purrr::partial(DiagrammeR::grViz, diagram = options$code)
  
  out <- do.call(dia_call, grviz_opt) 
  options$code <- out$x$diagram
  options$engine <- 'dot'
  options$echo <- FALSE
  knitr::knit_engines$get('dot')(options)
  # options$results <- 'asis'
  # knitr::engine_output(options, options$code, '')
}
knitr::knit_engines$set(diagrammer = my_fn$diagrammer_engine)

# Global theming for ggplots
my_fn$plot_theme <-
  theme(text = element_text('serif', size = 9), 
        plot.tag = element_text('serif', size = 9),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

na = \(x) sum(is.na(x))
```

```{r data, include=FALSE}
data_dir <- '../../data/cleaned/'
load(file.path(data_dir, 'data_input.Rdata'))
data_dirty <- readRDS(file.path(data_dir, 'data_dirty.RDS'))
setDT(data_dirty)
load(file.path(data_dir, '..', 'export', 'm3_summary.Rdata'))
load(file.path(data_dir, '..', 'export', 'm3_plot.Rdata'))
# m3 = readRDS(file.path(data_dir, '..', 'export', 'metrics', 'm3.RDS'))
m3 = new.env()
load(file.path(data_dir, '..', 'export', 'metrics', 'm3_calibROC.Rdata'), envir = m3)
```
\doublespacing
# Abstract {-}

## Background {-}

*Tuberculous meningitis* (TBM) is the one most lethal form of tuberculosis, especially for HIV-infected patients. There is no diagnostic model whereas current uniform case definition is based on expertise and comes with high uncertainty. We developed a model based on Latent class analysis (LCA) that provides a statistical scoring system estimating individual TBM risk amongst adults with suspected brain infection, taking into account the uncertainty of TBM diagnosis.

## Methods {-}

The model consisted of two components: one LCA that linked latent TBM status with three bacteria-detection confirmatory tests as manifest variables; and a prevalence model related relevant risk factors with TBM status via a logistic regression. The local dependence between manifest variables are corrected by latent bacillary burden, modelled via a linear mixed-effect model. Five model designs were considered and Bayesian approach was used to incorporates prior knowledge. Missing data was treated by their anticipated missing mechanisms. All considered models were validated using 5 repeated 20-fold cross validation.

## Results {-}

We estimated the odd ratios of TBM for several risk factors together with bacillary burden at diagnosis. Our selected model showed good AUC of 92.4%, 92.9%, and 96.1% against 3 confirmatory tests and good calibration (intercepts = 0.05, 0.1, 0.17 and slopes = 1.1, 1.2, 1.3). Against the hospital diagnosis made by physicians during follow-up, AUC was 93.9%, with calibration intercept of -0.01 and slope was 0.97. 

## Conclusion {-}

Compared to the uniform case definition, our model provides a well-calibrated scoring system with good correspondence to the current practice. Our model also gives an insight of mechanistic pathway from disease status to test results via disease bacillary burden at diagnosis time.

