---
title: 'A novel scoring system for Tuberculous Meningitis Diagnosis: a Bayesian Latent Class Model'
site: bookdown::bookdown_site
bibliography: includes/references.bib
csl: includes/clinical-infectious-diseases.csl
fontsize: 12pt
output:
  bookdown::pdf_book:
    base_format: rmarkdown::latex_document
    toc: no
    pandoc_args:
      - '--lua-filter=includes/scholarly-metadata.lua'
      - '--lua-filter=includes/author-info-blocks.lua'
      - '--standalone'
    latex_engine: xelatex
    highlight: pygments
    includes:
      in_header: includes/header.tex
---

```{r setup, include=FALSE}
options(tinytex.engine_args = '-shell-escape')
knitr::opts_chunk$set(echo = FALSE, dpi=300) #, dev = 'png'
flextable::set_flextable_defaults(fonts_ignore=TRUE, font.size=9, table.layout='fixed')
```

```{r library, include=FALSE}
library(gtsummary)
library(dplyr)
library(data.table)
library(magrittr, include.only = '%$%')
library(ggplot2)
library(patchwork)
library(officedown)

# All the misc functions go here.
my_fn <- new.env()

## Add labels accross a tbl
my_fn$add_labels <- \(.data, ...){
  cols <- list(...)
  for (i in seq_along(cols))
    attr(.data[[names(cols)[i]]], 'label') <- cols[[i]]
  .data
}

## Register a knit engine for stan code highlight
my_fn$stan_knitr_engine <- \(options){
  out <- c('\\begin{Shaded}','\\begin{minted}{stan}', options$code, '\\end{minted}','\\end{Shaded}')
  options$echo <- FALSE
  options$results <- "asis"
  knitr::engine_output(options, options$code, out)
}
knitr::knit_engines$set(stan_code = my_fn$stan_knitr_engine)

## Gist to number equations
## https://gist.github.com/mcanouil/eb75057432ff77846f4273d8808e615a
my_fn$is_docx_output <- function (fmt = knitr:::pandoc_to()) {
  if (length(fmt) == 0) {
    return(FALSE)
  } else {
    return(fmt == "docx")
  }
}

my_fn$numberEq.docx <- function (eq, lab, envir = docx.eqcounter) {
  assign(x = "counter", value = get(x = "counter", envir = envir)+1, envir = envir)
  assign(x = gsub("eq:", "", lab), value = get(x = "counter", envir = envir), envir = envir)
  lab <- get("counter", envir = envir)
  return(c('$$', eq, '\\;\\;\\;\\;(', lab, ')', '$$'))
}
my_fn$labEq.docx <- function (lab, envir = docx.eqcounter) {
  return(paste0('(', get(x = gsub("eq:", "", lab), envir = envir), ')'))
}

docx.eqcounter <- new.env()
docx.eqcounter$counter <- 0
  
# Register a knit engine for formula
my_fn$aligned_formula_knit <- \(options){
  out <- my_fn$numberEq.docx(options$code, lab = options$fml.lab)
  options$echo <- FALSE
  options$results <- "asis"
  knitr::engine_output(options, options$code, out)
}
knitr::knit_engines$set(aligned_formula = my_fn$aligned_formula_knit)

# Register diagrammer graphviz
my_fn$diagrammer_engine <- \(options){
  grviz_opt = formals(DiagrammeR::grViz)
  grviz_opt = grviz_opt[names(grviz_opt) != 'diagram']
  new_opt = options[names(options) %in% names(grviz_opt)]
  # not_grviz_opt = opt[!names(opt %in% c('diagram', name(grviz_opt)))]
  grviz_opt = modifyList(grviz_opt, new_opt)
  dia_call <- purrr::partial(DiagrammeR::grViz, diagram = options$code)
  
  out <- do.call(dia_call, grviz_opt) 
  options$code <- out$x$diagram
  options$engine <- 'dot'
  options$echo <- FALSE
  knitr::knit_engines$get('dot')(options)
  # options$results <- 'asis'
  # knitr::engine_output(options, options$code, '')
}
knitr::knit_engines$set(diagrammer = my_fn$diagrammer_engine)

# Global theming for ggplots
my_fn$plot_theme <-
  theme(text = element_text('serif', size = 9), 
        plot.tag = element_text('serif', size = 9),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

na = \(x) sum(is.na(x))
```

```{r data, include=FALSE}
data_dir <- '../../data/cleaned/'
load(file.path(data_dir, 'data_input.Rdata'))
data_dirty <- readRDS(file.path(data_dir, 'data_dirty.RDS'))
setDT(data_dirty)
load(file.path(data_dir, '..', 'export', 'm3_summary.Rdata'))
load(file.path(data_dir, '..', 'export', 'm3_plot.Rdata'))
# m3 = readRDS(file.path(data_dir, '..', 'export', 'metrics', 'm3.RDS'))
m3 = new.env()
load(file.path(data_dir, '..', 'export', 'metrics', 'm3_calibROC.Rdata'), envir = m3)
load(file.path(data_dir, '..', 'export', 'm3_plot.Rdata'), envir = m3)
m3$rocs = readRDS(file.path(data_dir, '..', 'export', 'metrics', 'm3_roc_iscb.RDS'))
m3$calib = readRDS(file.path(data_dir, '..', 'export', 'metrics', 'm3_calib_updated.RDS'))
```
\doublespacing
# Introduction

*Tuberculous meningitis (TBM)* is the most lethal form of tuberculosis. The diagnosis of TBM is challenging, as there is no gold standard. Results of methods to detect bacteria in cerebrospinal fluid (CSF) samples including Ziehl-Neelsen staining (Smear), GeneXpert MTB/RIF (Xpert), and culturing in Mycobacteria Growth Indicator Tube (MGIT) are only known after 1 days to 4 weeks. All lack sensitivity [@nhu2013; @chen2020]. In 2010, a uniform case definition was agreed upon amongst experts [@marais2010] but did not have a formal quantification and was not intended as a diagnostic tool. Given the deadly consequences, physicians usually incorporate their experience to make the decisions.

*Latent class analysis (LCA)* is a modelling technique that incorporates hidden traits and assigns a probability of traits based on binary manifest variables. Auxiliary variables can be added to guide the task. Our study aim to use those auxiliary variables and build a scoring system that estimates the probability of TBM based on data with uncertain TBM status.

# Methods

We used data from an observational study of brain infection conducted in Ho Chi Minh City, Vietnam. Our model consists of two components. The basic component links 3 manifest variables ($Y$) to the latent TBM status ($C$). Local dependence between tests was addressed by a representation of bacillary burden (*B*), modelled as a combination of a random effect (*r*) [@qu1996] and covariates (*V*). The other component is a logistic regression prevalence model relating TBM status to the risk factors (*X*) which were transformed to the same scale [@gelman2008]. Except for the false positive rate ($\zeta^{(t)}_0$) for which we chose a strong prior based on literatures [@nhu2013], we used weakly informative priors for the parameters.
$$
Y^{(t)} \sim Bernoulli(\rho^{(t)})
$$
$$
\rho^{(t)} 
=\sum_{c=0}^1Pr(C=c) * Pr(Y^{(t)}=1|C=c)
= \theta * logit^{-1}(\zeta_1^{(t)} + \beta^{(t)}B) + (1 - \theta) * logit^{-1}(\zeta_0^{(t)})
$$
$$
\begin{aligned}
\theta &= logit^{-1}(X^T \alpha) &B &= V^T \gamma + r \\
\zeta^{(t)}_1 &\sim Logistic(0,0.35) 
&r &\sim \mathcal{N}(0, 1) 
&{\alpha,\ \beta^{(t)}} &\sim t_5(0, s) &\gamma \sim t_5(0, \frac{s}{\overline{\beta^{(t)}}})
&\ &s \sim \mathcal{N}(0, 1.5)
\end{aligned}
$$
*where $\theta$ is the TBM probability ($Pr(C=1)$), $\rho^{(t)}$ is the probability of a positive test (t with t = {Smear, MGIT, Xpert}) whose observed values are $Y^{(t)}$; $\alpha$, $\beta$, $\gamma$ are the coefficients or $X$, $B$, and $V$. To avoid sign switching, $\beta$ is constrained to be positive.*

We considered 4 model designs. Model 1 was similar to M1 in [@schumacher2016], as no bacillary burden was included ($\beta^{(t)} = 0$). In model 2 we assumed the impact of bacillary burden on tests to be equal ($\beta^{Smear} = \beta^{MGIT} = \beta^{Xpert}$), which was lifted in model 3. Model 4 was similar to M4 in [@schumacher2016], as we moved $V$ out of the bacillary burden components, directly linking it to $\rho^{(t)}$. We assessed local independence via residual pairwise correlation between 3 tests [@qu1996]. _Expected log point-wise predictive density (elpd)_ [@vehtari2016] was used for model selection. We assessed the calibration [@VanCalster2019], receiver operating characteristic (ROC) curves and area under the curve (AUC) of our models against 3 manifest variables. We also pseudo-validated our prevalence model against the "hospital diagnosis" made by the physicians at the end of the follow-up. Five repeated 20-fold cross validations were used in all validations. Missing covariates were treated case by case according to their missing mechanisms (Table \@ref(tab:missing-handling)). All analyses were performed using the statistical package \textsf{R}, version 4.1.2 [@rcoreteam] and \textsf{Stan}, version 2.27 [@stan-doc].<!-- All code were published on [project's github repo](github.com/TBM-LCA).-->

# Results

```{r est_fun, include=FALSE}
est = \(name, tab, digits=2, transformation = I, ci.sep = ', ', post.fn = I, pre.fn=I) {
  m <- c(pre.fn(tab[name, 'mean']), pre.fn(tab[name, '2.5%']), pre.fn(tab[name, '97.5%'])) |> transformation() 
  c(m[1]|> formatC(digits=digits, format='f') |> post.fn(),
    # m[2]|> formatC(digits=digits, format='f') |> post.fn(), 
    paste(min(m[2:3]) |> formatC(digits=digits, format='f') |> post.fn(), 
          max(m[2:3]) |> formatC(digits=digits, format='f') |> post.fn(), sep=ci.sep))
}

spc.est = purrr::partial(est,
                         transformation=\(x) 100*(1-x), ci.sep=' - ', digits=1,
                         post.fn = \(x) paste0(x, '%'))
sen.est = purrr::partial(est, transformation=\(x) 100*x, ci.sep=' - ', digits=1,
                         post.fn = \(x) paste0(x, '%'))

spc.est2 = purrr::partial(est,
                         transformation=\(x)  100*(1-x), ci.sep=', ', digits=1)
sen.est2 = purrr::partial(est, transformation=\(x) 100*x, ci.sep=', ', digits=1)

sen = \(name, hiv=NULL, fn=sen.est, sep = ' ') {
  name = paste0(name, '[2]')
  out = do.call(sprintf, as.list(c('%s (%s)', fn(name, 
                                                tab = if (!length(hiv)) p_summary else if(hiv) p_hiv_pos_summary else p_hiv_neg_summary))))
  gsub('\\s\\(', paste0(sep, '('), out)
}

spc = \(name, hiv=NULL, fn=spc.est, sep = ' ') {
  name = paste0(name, '[1]')
  out = do.call(sprintf, as.list(c('%s (%s)', fn(name,
                                                tab = if (!length(hiv)) p_summary else if(hiv) p_hiv_pos_summary else p_hiv_neg_summary))))
  gsub('\\s\\(', paste0(sep, '('), out)
}
```

Estimated coefficients are shown in Figure \@ref(fig:fig-est). Patients with HIV, Miliary TB, and long symptom duration had higher risk of TBM. HIV positivity and higher CSF protein were associated with higher bacillary burden. Although higher CSF lymphocyte relates to TBM, it was associated with a lower bacillary burden. Except for model 1, all residual pairwise correlations fluctuated around 0. Model 3 was selected with highest elpd. It was well-calibrated and had high discriminative value against 3 tests (AUC = 92.4 - 96.1%) (Figure \@ref(fig:model-metrics)). Against hospital diagnosis, we also saw good calibration and high AUC (93.9%). 

# Discussions

We used latent class analysis to develop a TBM diagnostic model, included random effects to obtain local independence between tests. Different to a previous study [@schumacher2016], all factors (*V*) were wrapped inside the bacillary burden (*B*); this gained us some degrees of freedom and provided an insight into the mechanistic pathway from disease status to tests. The model had good calibration against the confirmatory tests. It also had good calibration against hospital diagnosis, suggesting a good correspondence between our model and clinical judgement.

Our study has some limitations. For missing data we had to make assumptions that can bias the results. We also had to depend on another hypothesized variable (bacillary burden) that cannot be validated. Further research in this field is thereby needed. 

# Conclusion

Our model provides a well-calibrated and highly predictive diagnostic model for TBM diagnosis. Until a true gold standard is developed, our model can be an effective diagnostic tool in both clinical practice and research for TBM patients.

# References

::: {#refs}
:::

\newpage

```{r skeleton-model, eval=FALSE, fig.align="center", fig.cap="Basic model design: TBM risk factors predict individual's TBM status. Unknown TBM status is linked with test results. The probability of a positive test depends on bacillary burden. Test positive probabilities are for demonstration only and do not correspond to the real one.", fig.id='skeleton-model', fig.show='hold', include=FALSE, out.height='10cm', out.width="100%"}

fig_svg <- cowplot::ggdraw() + 
  cowplot::draw_image(magick::image_read_svg("includes/classicLCA.svg", width=212*5, height=159*5))
plot(fig_svg)
```
\singlespacing
```{r missing-handling, tab.cap="Missing values rationales and handling", tab.id="missing-handling", out.width='100%'}
na = \(x) sum(is.na(x))
pop = data_dirty[,(wrong_name%in%F) &
    !((Volume<3)%in%TRUE) &
    !(csf_mgit_contaminated %in% T) ]
data_dirty$csf_neutro <- data_dirty$NEUPER * data_dirty$csf_wbc / 100
data_dirty |>
  dplyr::filter(pop) %$%
  # dplyr::filter(!(wrong_name %in% TRUE) & !(csf_mgit_contaminated %in% TRUE)) %$%
  tibble::tribble(
     ~ 'Variable'           , ~ 'N\n missing'     ,  ~ 'Expected Reason of Missingness'                      , ~ 'Mechanism',        ~ 'Handling method' ,
    'ZN Smear'              , na(csf_smear)       , 'Not suspected TBM'                                      , 'MNAR'               , 'Set = 0'          , 
    'MGIT'                  , na(csf_mgit)        , 'Not suspected TBM'                                      , 'MNAR'               , 'Set = 0'          , 
    'Xpert'                 , na(csf_xpert)       , 'Not suspected TBM'                                      , 'MNAR'               , 'Set = 0'          , 
    'HIV Status'            , na(hiv_stat)        , 'Not suspected HIV'                                     , 'MNAR'           , 'Set = 0'       ,
    'TB-suggested symptoms' , na(clin_symptoms)   , 'Unmeasured / Unnoticed / Unconscious'                   , 'MAR'       , 'Imputation'       ,
    'Focal neuro-deficit'   , na(clin_motor_palsy), 'Unconscious / Pre-existed'                              , 'MAR'           , 'Imputation'       ,
    'Glasgow Coma Score'    , na(clin_gcs)        , 'Ventilated (GCSV) / Input error'       , 'MAR'                , 'Imputation'       ,
    'Illness days'          , na(clin_illness_day), 'Patients forget / Unconscious'                          , 'MAR'                , 'Imputation'       ,
    # 'Blood Lymphocyte'      , na(LYMP)            , 'Unmeasured (premature death)'                           , 'MAR'                , 'Imputation'       ,
    # 'Blood Neutrophil'      , na(NEUTRO)          , 'Unmeasured (premature death)'                           , 'MAR'                , 'Imputation'       ,
    'Blood Glucose'         , na(BLDGLU)          , 'Most likely input error / Unmeasured (premature death)', 'MAR/MCAR'           , 'Imputation'       ,
    'CSF glucose'           , na(csf_glucose)     , 'Unmeasured (premature death)'                           , 'MAR/MCAR'           , 'Imputation'       ,
    'CSF lymphocyte count'  , na(csf_lympho)      , 'Very low or zero / Input error / Unmeasured (premature death)', 'MNAR/MAR'           , 'Manually set/Imputation',
    'CSF WBC count'  , na(csf_wbc)      , 'Very low or zero / Input error / Unmeasured (premature death)', 'MNAR/MAR'           , 'Manually set/Imputation',
    'CSF protein'           , na(csf_protein)     , 'Data input error / Unmeasured (premature death)'       , 'MAR/MCAR'           , 'Imputation'       ,
    'CSF lactate'           , na(csf_lactate)     , 'Unmeasured (premature death)'                              , 'MAR/MCAR'           , 'Imputation' ,    
    # 'CSF RBC count'         , na(csf_rbc)         , 'Zero cell count'                                        , 'MNAR'               , 'Set = 0',
    # 'CSF eosinophil count'         , na(csf_rbc)         , 'Zero cell count'                                        , 'MNAR'               , 'Set = 0',
    'Cryptococcal test'         , sum(!(CRYTO=='NOT DONE'&INDIAINK=='NOT DONE')%in%F) , 'Not suspected cryptococcal meningitis'                                        , 'MNAR'               , 'Set = 0',
    'Gram stain'         , sum(is.na(GRAM) | GRAM=='NOT DONE') , 'Not suspected bacterial meningitis'                                        , 'MNAR'               , 'Set = 0'
  ) |>
  flextable::flextable() |>
  # flextable::set_caption('Rationale and method of missing values handling') |>
  flextable::width(width = 1.25) |>
  # flextable::footnote(i = c(11,12),j = 1,inline=FALSE,
  #                     value=flextable::as_paragraph(rep('CSF Lymphocyte count was calculated by CSF WBC count * Lymphocyte% / 100; if very low, then one Lymphocyte%, our Neutrophils% had value, the other were left missing, implying a 0 or left over')),
  #                     part = 'body') |>
  flextable::add_footer_row(values = 'MAR: missing at random; MNAR: missing not at random; MCAR: missing completely at random', colwidths = 5) |>
  flextable::merge_v(part='footer') |>
  flextable::theme_vanilla() |>
  flextable::bold(bold=FALSE, part='footer') |>
  flextable::italic(italic=TRUE, part='footer')
```

```{r mv-posterior, tab.cap="Posterior estimates of test sensitivities and specificities, for overall population, and stratified by HIV", tab.id="mv-posterior", include=FALSE}
# Use %
pr_neg <- matrix(c(
  c('Test', rep(c('Specificity (%)'),1), rep(c('Sensitivity (%)'),1)),
  # c('Test', rep(c('Estimates'),2)),
  c('Smear',
    spc("z_Smear", hiv=FALSE, fn=spc.est2, sep='\n'),
    sen("z_Smear", hiv=FALSE, fn=sen.est2, sep='\n')),
  c('MGIT',
    spc("z_Mgit", hiv=FALSE, fn=spc.est2, sep='\n'),
    sen("z_Mgit", hiv=FALSE, fn=sen.est2, sep='\n')),
  c('Xpert',
    spc("z_Xpert", hiv=FALSE, fn=spc.est2, sep='\n'),
    sen("z_Xpert", hiv=FALSE, fn=sen.est2, sep='\n'))
), byrow=T, nrow=4)

pr_pos <- matrix(c(
  c('Test', rep(c('Specificity (%)'),1), rep(c('Sensitivity (%)'),1)),
  c('Smear',
    spc("z_Smear", hiv=TRUE, fn=spc.est2, sep='\n'),
    sen("z_Smear", hiv=TRUE, fn=sen.est2, sep='\n')),
  c('MGIT',
    spc("z_Mgit", hiv=TRUE, fn=spc.est2, sep='\n'),
    sen("z_Mgit", hiv=TRUE, fn=sen.est2, sep='\n')),
  c('Xpert',
    spc("z_Xpert", hiv=TRUE, fn=spc.est2, sep='\n'),
    sen("z_Xpert", hiv=TRUE, fn=sen.est2, sep='\n'))
), byrow=T, nrow=4)

pr <- matrix(c(
  c('Test', rep(c('Specificity (%)'),1), rep(c('Sensitivity (%)'),1)),
  c('Smear',
    spc("z_Smear", hiv=NULL, fn=spc.est2, sep='\n'),
    sen("z_Smear", hiv=NULL, fn=sen.est2, sep='\n')),
  c('MGIT',
    spc("z_Mgit", hiv=NULL, fn=spc.est2, sep='\n'),
    sen("z_Mgit", hiv=NULL, fn=sen.est2, sep='\n')),
  c('Xpert',
    spc("z_Xpert", hiv=NULL, fn=spc.est2, sep='\n'),
    sen("z_Xpert", hiv=NULL, fn=sen.est2, sep='\n'))
), byrow=T, nrow=4)

huxtable::as_hux(cbind(pr, pr_neg[,-1], pr_pos[,-1])) |>
  huxtable::as_flextable() |>
  flextable::add_header_row(values=c("", rep("Overall",2), rep("HIV (-)",2), rep("HIV (+)",2))) |>
  flextable::merge_h(i=1, part='header') |>
  flextable::width(width = .8) |>
  flextable::bold(i=1) |>
  flextable::theme_vanilla() |>
  flextable::align(i=1, align='center') |>
  flextable::align(i=1, align='center', part='header') |>
  flextable::add_footer_row(top=F, values="Values are in the form Mean (95% Credible interval)", colwidths = 7) -> restab
# restab_raster = flextable::as_raster(restab |> flextable::width(width=1.2), zoom=1.2)
restab
```

<!-- \newpage -->
<!-- \blandscape -->

```{r fig-est, out.width="90%", fig.dim=c(8, 10), fig.align="center", fig.cap="A: Estimates of TBM odd ratios in prevalence model for different covariates, excluding WBC count; B: Estimates of TBM odd ratio for number of CSF WBC count; C: Estimates of factors impacting individual's bacillary burden, given they have TBM. In A and C, dot, thick lines, and thin lines are posterior expected values, 50\\% and 95\\% credible intervals. In B: the blue line is expected odds ratio, the inner ribbon and outer ribbon are 50\\% and 95\\% credible intervals.", fig.id="fig-est", warning=FALSE, message=FALSE}

library(patchwork)

# rotate xlab 45deg
# log10(eos) > 0 wrong -> eos > 0
# odd ratio
# C: standardised bacillary burden #explain in text
y_labs = m3$a_plot$scales$get_scales('y')$labels
y_labs[3] = "*log<sub>10</sub>* (CSF RBC)"
y_labs[4] = "*log<sub>10</sub>* (CSF Eosinophil)" 
y_labs[5] = 'CSF Eosinophil > 0'
y_labs[6] = "*log<sub>10</sub>* (CSF Lymphocyte)"

y_labs2 = m3$b_plot$scales$get_scales('y')$labels
y_labs2[1] = "*log<sub>10</sub>* (CSF WBC)"  
y_labs2[2] = "*log<sub>10</sub>* (CSF Lymphocyte)"

ab_plot = ((m3$a_plot |> td.misc::change_ylabs(rev(y_labs), top_down = TRUE)) + ggtitle("Prevalence model") + 
             scale_x_continuous(
               name="TBM odd ratio",
               breaks = m3$a_plot$scales$get_scales('x')$breaks,
               labels = m3$a_plot$scales$get_scales('x')$labels) + 
             theme(axis.text.y = ggtext::element_markdown())) / 
  (m3$wbc_plot + theme(axis.title.y=element_text(size=8, face='bold',vjust=-45, color=grey(.3))) +
     scale_y_continuous(
    name = 'TBM odd ratio',
    breaks = log(10^c(-24,-20, -16,-12, -8,-4,-2, -1, 0)),
    labels = c('1e-24','1e-20','1e-16','1e-12','1e-8', '1e-4', 0.01, .1, 1)
  )) /
  ((m3$b_plot |> td.misc::change_ylabs(rev(y_labs2), top_down = TRUE)) + 
     ggtitle("Bacillary burden model") + 
     theme(axis.text.y = ggtext::element_markdown()) + 
     scale_x_continuous(name='Standardised bacillary burden')) /
  # grid::rasterGrob(restab_raster) + 
  patchwork::plot_layout(ncol=1, nrow=3, 
                         heights = c(6,2,2)) + 
  plot_annotation(tag_levels="A")

#ggtitle("CSF WBC count") + theme(title = element_text(face='bold',size=6), title.po

# plt = patchwork::wrap_plots(ab_plot, m3$plot.multggplotGrob(m3$rocs)) + plot_annotation(tag_levels="A")
# plot(plt)
ab_plot
```

```{r model-metrics, fig.align="center", fig.cap='Model performance. A: ROC and AUC against confirmatory test results; B: calibration against confimatory test results; C: ROC and AUC against hospital diagnosis; D: Calibration against hospital diagnosis. For calibration plots, “observed” probabilities are defined as values smoothed by a loess fit against observed binary events, in which the translucent ribbons represent the level of uncertainty of the smoothing. For ROC plots, AUC values are presented as \\emph{“average (min - max over 5 repetitions of cross-validation)”}. In all plots, grey lines are fitted curves from each 20-fold cross validation and coloured lines represents their average.' , fig.dim=c(9, 12), fig.show='hold', message=FALSE, warning=FALSE, out.width='80%', fig.id="model-metrics"}

design <- "
1111
2222
3333
"
#remove youden index
#look for the scale of calib
#remove 

roc_y <- do.call(gridExtra::arrangeGrob, m3$rocs$Y)
m3$calib$Y2[[3]] = m3$calib$Y2[[3]] + theme(axis.title.y = element_blank())
(wrap_elements(roc_y) / wrap_elements(m3$calib$Y2)) /
  (wrap_elements(m3$rocs$C) + wrap_elements(m3$calib$C)) + 
  plot_layout(design = design, 
              # widths=c(1,33,33,33),
              heights=c(2,2,3)) +
  plot_annotation(tag_levels = 'A')


```

<!-- \elandscape -->
