---
title: "Update of TBM Latent Class Analysis"
author: Trinh Dong Huu Khanh
date: "4/12/2020"
output: |
  bookdown::pdf_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(data.table)
```

# Overview

## Objectives

- Give an improved score table based on Statistics replacing the current, partially expertise-based, TBM definition Consensus Score (2010)
- Give a predicted probability of TBM at roughly admission time...
- ...And after confirmation test results (mainly all negative)

## Concept and Terms

- Latent Class Analysis = Finite Bernoulli Mixture Model
- Manifest variable = (Confirmation) Test = Indicator
- Predictor = Co-variate
- Prevalence = Theta = Latent class = Latent variable

```{dot echo=FALSE, out.width="80%", out.height="40%", fig.align="center", fig.cap="Model Concept"}
digraph Model_Concept{
  rankdir = LR;
  resolution=300;
  compound=true;
  
  subgraph cluster_predictor{
    color=gray;
    node [style=filled];
    predictor1; predictor2; predictor3;
  }
  
  subgraph cluster_manifest{
    node [style=filled];
    manifest1; manifest2; manifest3;
  }
  
    
  predictor1[label="Predictor 1", shape="box"];
  predictor2[label="Predictor 2", shape="box"];
  predictor3[label="Predictor 3", shape="box"];
  manifest1[label="Manifest Variable 1", shape="box"];
  manifest2[label="Manifest Variable 2", shape="box"];
  manifest3[label="Manifest Variable 3", shape="box"];
  latent[label="Class", shape="oval"];
  
  {predictor1 predictor2 predictor3} -> latent -> {manifest1 manifest2 manifest3};
  predictor1 -> manifest1
}
```


# Current approach

## Core model

I use a similar set of clinical and laboratorial signs and symptoms to the TBM definition score as **predictors**. Weakly $\mathrm{t}$ distribution are used.

Continuous predictors are transformed to have a symmetric distribution.

$$
\begin{aligned}
  nu \sim \mathcal{\Gamma}(2,0.1) \\
  a_0 \sim \mathrm{t}(\nu, 0, 2.5) \\
  a \sim \mathrm{t}(\nu, 0, 1)
\end{aligned}
$$

**Three** confirmation tests are chosen as manifest variables. Their **False positive rate (FPR)** and **True positive rate (TPF)** are assumed to follow a *Normal* distribution on the logit scale. 

Informative priors are imposed on **FPR** according to priors knowledge and experience (thanks to Joe and Julie from TB). For **TPR**, a weakly informative was used with sufficient collapse toward .5 after transformation back to normal scale.

$$
\begin{aligned}
  TPR \sim \mathcal{N}(0, .6) \\
  \\
  FPR_{Smear} \sim \mathcal{N}(logit(.001),\ \ .82)\\
  FPR_{Mgit} \sim \mathcal{N}(logit(.001),\ \ .82)\\
  FPR_{Xpert} \sim \mathcal{N}(logit(.005),\ 1.59)
\end{aligned}
$$
**Bacillary burden** was assumed to be a random variable whose mean depends on **HIV** status.

$$
\begin{aligned}
  b_{RE} \sim \mathrm{t}(\nu, 0, 1)\\
  b_{HIV} \sim \mathrm{t}(\nu, 0, 1)\\
  RE \sim \mathcal{N}(0, 1)\\
  bac\_load = b_{RE} * RE + b_{HIV} * HIV
\end{aligned}
$$
```{dot CoreModel, echo=FALSE, out.width="80%", fig.align="center", fig.cap="Core model"}
digraph Model_Core{
  rankdir = LR;
  resolution=500;
  compound=true;
  
  subgraph cluster_predictor{
    color=gray;
    node [style=filled];
    HIV; Symptoms; Motor; Nerve; PulTB; MilTB; Age; TBDays; BldGlu; CsfGlu; GluRatio; CsfLymp; CsfPro; CsfLac; GCS;
    label="Predictors"
  }
  
  subgraph cluster_manifest{
    color=gray;
    Xpert[shape="box"]; Mgit[shape="box"]; Smear[shape="box"];
    label="Confirmation Tests"
  }
  
  subgraph cluster_latent{
    color=none;
    latent;bacload;
  }
  
  HIV[label="HIV Status", shape="box"];
  Symptoms[label="TB symptoms", shape="box"];
  Motor[label="Motor palsy", shape="box"]; 
  Nerve[label="Cranial palsy", shape="box"]; 
  PulTB[label="X-ray Pulmonary TB", shape="box"];
  MilTB[label="X-ray Miliary TB", shape="box"];
  Age[label="Age", shape="box"];
  TBDays[label="Illness days", shape="box"]; 
  BldGlu[label="Blood Glucose", shape="box"];
  CsfGlu[label="CSF Glucose", shape="box"]; 
  GluRatio[label="Glucose Ratio", shape="box"];
  CsfLymp[label="CSF Lympho", shape="box"];
  CsfPro[label="CSF Protein", shape="box"];
  CsfLac[label="CSF Lactate", shape="box"];
  GCS[label="Glasgow Coma Score", shape="box"];
  
  latent[label="TBM Status", shape="oval"];
  
  bacload[label="Bacillary Burden", shape="oval", style="filled", color="grey"];
  
  # {HIV  Symptoms  Motor  Nerve  PulTB  MilTB  Age  TBDays  BldGlu  CsfGlu  GluRatio  CsfLymp  CsfPro  CsfLac  GCS} -> latent -> {Xpert Mgit Smear};
  Age -> latent[ltail=cluster_predictor]
  latent -> Mgit[lhead=cluster_manifest];
  
  HIV -> bacload; 
  bacload -> {Mgit}[lhead=cluster_manifest];
}

```

## Imputation model

4 cases with missing data in confirmation tests are removed. 

Missing data in predictors are assumed MAR and imputed within the model using _Stan_.
  
  - _HIV_ is imputed using **logistic regression** and have the probabilities marginalised.
  
$$HIV_{a_0} \sim \mathrm{t}(\nu, 0, 1) \\$$
  - Those which are compound of several other binary variables, such as _TB Symptoms_ = _Weight Loss_ || _Night Sweats_ || _Coughing_, have there respective compartments imputed and combined: _TB symptoms_, _Motor Palsy_. These compartments are in turn imputed using a **multivariate probit regression**.
  
$$
\begin{aligned}
  L_{Omega_{cs}} \sim LKJCorrelationCholesky(4) \\
  cs_{a_0} \sim \mathrm{t}(\nu, 0, 1) \\
  cs_a \sim \mathrm{t}(\nu, 0, 1) \\
  cs_z \sim \mathcal{N}(cs_{a_0} + cs_a^{(1)}*HIV + cs_a^{(2)}*TB\_Day, L_{Omega_{cs}});
\end{aligned}
$$
  
  - Clinical continuous variables, after transformed, are imputed using **linear regression** corrected for **HIV Status**: _Age_, _TB Days_, _GCS_. 
  
  
$$
\begin{aligned}
age_{a_0} \sim \mathrm{t}(\nu, 0, 1)\\
age_{a} \sim \mathrm{t}(\nu, 0, 1)\\
age \sim \mathcal{N}(age_{a_0} + age_a*HIV, age_{\sigma})
\end{aligned}
$$
  
  - Lab values are imputed toegether using **seemingly unrelated regression**: _Blood Glucose_, _CSF Glucose_, _CSF Lymphocyte count_, _CSF Protein_, _CSF Lactate_. Formulation similar to those of **multivariate probit regression**.
  
  - _GCS_, integer, which is a linear sum of _GCSV_, _GCSM_, _GCSE_, have its only missing value imputed as a continuous variable.
  
$$
\begin{aligned}
GCSV_{a_0} \sim \mathrm{t}(\nu, 0, 1)\\
GCSV_{a} \sim \mathrm{t}(\nu, 0, 1)\\
GCSV_\sigma \sim \mathcal{N}(0,1)\\
GCSV \sim \mathcal{N}(GCSV_{a_0} *\begin{bmatrix}
    GCSE\\
    GCSM
  \end{bmatrix}, 
  GCSV_\sigma) \\
GCS = GCSV + GCSM + GCSE
\end{aligned}
$$

```{dot ImpModel, echo=FALSE, out.width="100%", out.height="100%", fig.align="center", fig.cap="Imputation model"}
digraph Model_Imp{
  rankdir = LR;
  resolution=500;
  compound=true;
  
  subgraph cluster_clinsymp{
    Cough[label="Coughing"];
    Nsweats[label="Night Sweats"];
    WLoss[label="Weight Loss"];
  }
  
  subgraph cluster_motor{
    Hemi[label="Hemiplegia"];
    Para[label="Paraplegia"];
    Tetra[label="Tetraplegia"];
  }
  
  subgraph cluster_gcs{
    GCSV; GCSM; GCSE;
  }
  
    
  HIV[label="HIV Status", shape="box"];
  Symptoms[label="TB symptoms", shape="box"];
  Motor[label="Motor palsy", shape="box"];
  Age[label="Age", shape="box"];
  TBDays[label="Illness days", shape="box"]; 
  GluRatio[label="Glucose Ratio", shape="box"];
  BldGlu[label="Blood Glucose", shape="box"];
  CsfGlu[label="CSF Glucose", shape="box"]; 
  CsfLymp[label="CSF Lympho", shape="box"];
  CsfPro[label="CSF Protein", shape="box"];
  CsfLac[label="CSF Lactate", shape="box"];
  GCS[label="Glasgow Coma Score", shape="box"];
  
  HIV -> {Symptoms Age TBDays};
  WLoss -> Symptoms[ltail=cluster_clinsymp];
  Tetra -> Motor[ltail=cluster_motor];
  GCSM -> GCS[ltail=cluster_gcs];
  TBDays -> {Symptoms Age};
  
  
  Cough -> Nsweats[dir=both];
  Nsweats -> WLoss[dir=both];
  WLoss -> Cough[dir=both];
  Hemi -> Para[dir=both];
  Para -> Tetra[dir=both]; 
  Tetra -> Hemi[dir=both];
  BldGlu -> CsfGlu[dir=both];
  CsfGlu -> CsfLymp[dir=both];
  CsfLymp -> CsfPro[dir=both];
  CsfPro -> CsfLac[dir=both];
  CsfLac -> BldGlu[dir=both];
  {BldGlu CsfGlu} -> GluRatio;

  GCSV -> GCSE[dir=both];
  GCSE -> GCSM[dir=both];
  GCSM -> GCSV[dir=both];
}  
```

## Current Results

Below are current fit results for model **m2c** (model with Random Effects and GCS as continuous predictor), **m2cp** (GCS with quadratic effect), and **m2d** (GCS as dichotomous predictor, as used in the original definition score).

```{r load_files, message=FALSE}
library(ggplot2); library(bayesplot); library(patchwork); library(magrittr)
m2c_result <- readRDS("../outputs/m2c.RDS")
m2d_result <- readRDS("../outputs/m2d.RDS")
m2cp_result <- readRDS("../outputs/m2cp.RDS")
m2c_draw <- m2c_result$draws(c("a0", "a", "b_HIV", "b_RE", "z_Smear", "z_Mgit", "z_Xpert", "HIV_a0", "X"))
m2cp_draw <- m2cp_result$draws(c("a0", "a", "b_HIV", "b_RE", "z_Smear", "z_Mgit", "z_Xpert", "HIV_a0", "X"))
m2d_draw <- m2d_result$draws(c("a0", "a", "b_HIV", "b_RE", "z_Smear", "z_Mgit", "z_Xpert", "HIV_a0", "X"))

source("../r/include/functions.R")
```

```{r plot_result_a, message=FALSE, warning=FALSE, fig.asp=.8, fig.cap="Coefs in m2c and m2d"}
coef_plot.m2c <- bayesplot::mcmc_intervals(m2c_draw, 
                          pars=dplyr::vars(dplyr::starts_with("a")), 
                          prob_outer = .95,
                          point_est = "mean") %>% 
  change_ylabs("Intercept","HIV", "Symptoms", "Motor", "Nerve", "PulTB", "MilTB", 
              "Age", "TBDays", "sqrt(BldGlu)", "sqrt(CsfGlu)", "log2(CsfLymp)",
              "log2(CsfPro)", "log2(CsfLac)", "sqrt(GluRatio)", "log2(GCS)")

coef_plot.m2cp <- bayesplot::mcmc_intervals(m2cp_draw, 
                          pars=dplyr::vars(dplyr::starts_with("a")), 
                          prob_outer = .95,
                          point_est = "mean") %>% 
  change_ylabs("Intercept","HIV", "Symptoms", "Motor", "Nerve", "PulTB", "MilTB", 
              "Age", "TBDays", "sqrt(BldGlu)", "sqrt(CsfGlu)", "log2(CsfLymp)",
              "log2(CsfPro)", "log2(CsfLac)", "sqrt(GluRatio)", "log2(GCS)", "log2(GCS)^2")

coef_plot.m2d <- bayesplot::mcmc_intervals(m2d_draw, 
                          pars=dplyr::vars(dplyr::starts_with("a")), 
                          prob_outer = .95,
                          point_est = "mean") %>% 
  change_ylabs("Intercept","HIV", "Symptoms", "Motor", "Nerve", "PulTB", "MilTB"," 1(GCS<15)", "Age", "TBDays", "sqrt(BldGlu)", "sqrt(CsfGlu)", "log2(CsfLymp)", "log2(CsfPro)", "log2(CsfLac)", "sqrt(GluRatio)")

((coef_plot.m2c + ggtitle("m2c")) + (coef_plot.m2d + ggtitle("m2d")))
```

```{r plot_result_a2, message=FALSE, warning=FALSE, fig.cap="Coefs in m2c and m2cp"}
(coef_plot.m2c + ggtitle("m2c")) + (coef_plot.m2cp + ggtitle("m2cp"))
```

The estimation of sensitivities and specificities are similar for all three models.

```{r plot_result_FPR, message=FALSE, warning=FALSE, fig.cap=paste("False positive rates")}
FPR_plot_m2c_z <- bayesplot::mcmc_areas(m2c_draw, pars=dplyr::vars(dplyr::matches("^z.+[1]")), prob = .5) %>% change_ylabs("logit(FPR_Smear)", "logit(FPR_Mgit)", "logit(FPR_Xpert)")
FPR_plot_m2c_q <- bayesplot::mcmc_areas(m2c_draw, pars=dplyr::vars(dplyr::matches("^z.+[1]")), prob = .5, transformations = nimble::ilogit) %>% change_ylabs("FPR_Smear", "FPR_Mgit", "FPR_Xpert")

(FPR_plot_m2c_z + ggtitle("Logit scale")) + (FPR_plot_m2c_q + ggtitle("Normal scale"))
```

```{r plot_result_TPR, message=FALSE, warning=FALSE, fig.cap=paste("True positive rates") }
TPR_plot_m2c_z <- bayesplot::mcmc_areas(m2c_draw, pars=dplyr::vars(dplyr::matches("^z.+[2]")), prob = .5) %>% change_ylabs("logit(FPR_Smear)", "logit(FPR_Mgit)", "logit(FPR_Xpert)")
TPR_plot_m2c_q <- bayesplot::mcmc_areas(m2c_draw, pars=dplyr::vars(dplyr::matches("^z.+[2]")), prob = .5, transformations = nimble::ilogit) %>% change_ylabs("FPR_Smear", "FPR_Mgit", "FPR_Xpert")

(TPR_plot_m2c_z + ggtitle("Logit scale")) + (TPR_plot_m2c_q + ggtitle("Normal scale"))
```

## Problems

- Very complicated imputation model, especially with **discrete variables** $\Rightarrow$ Prone to errors. 
- Does illness days capture the effect from HIV? Should I include both?
- Problems with **Multivariate normal distribution** when doing partial prediction (i.e. imputation of partially missing combinations). **multi\_normal\_(cholesky)\_rng** in _Stan_ yields a whole vector by default $\Rightarrow$ Has to do manually with conditional distribution. Sometimes, case by case.
- More (missing) problems with new data??
- Switch to another language? [Pyro](http://pyro.ai/) seems to be a good replacement?
```{r data}
data_19EI <- readRDS('../data/cleaned/data_19EI.RDS')
data_19EI <- na.omit(data_19EI, cols = c('csf_smear', 'csf_mgit', 'csf_xpert'))
dt_filter <- data_19EI[is.na(csf_glucose) | is.na(BLDGLU) | is.na(csf_lympho) | is.na(csf_protein) | is.na(csf_lactate), .(USUBJID, bld_glucose=BLDGLU, csf_glucose, csf_lympho, csf_protein, csf_lactate)]
# gt::tab_header(gt::gt(dt_filter), title="Missing CSF lab results")
ht <- huxtable::set_all_borders(huxtable::set_caption(huxtable::as_hux(dt_filter), "Subset of individuals with missing CSF lab results"))
huxtable::caption_pos(ht) <- "bottom"
ht
```